---
title: "Equity Portfolio Decision Making"
author: Daniel Rabi
date: April 9, 2026
toc: true
toc_float: true
backgroundcolor: lightblue
format: html
editor: visual
embed-resources: true
bibliography: references/references.bib
csl: references/apa.csl
output-file: index
nocite: "@YahooFinance, @Gratton_2025, @Dolan_2025, @Kuepper_2025"

echo: FALSE
message: FALSE
warning: FALSE
---

![[@Zerdzicki_2026]](images/into_picture.jpg){fig-align="center" width="600"}

Have you ever wondered how to decide whether you should add equity to your existing portfolio?

This report evaluates whether adding TSLA, ENB, AMZN, LMT, or BWXT to an existing portfolio improves risk-adjusted returns, using machine learning forecasts and news sentiment.

# The Portfolio

This project's portfolio fairly diversified, unoptimized, existing portfolio that contains the following companies:

```{r, message=FALSE, warning=FALSE}
library(gt)

dplyr::tibble(
  Equity = c(
    "NVIDIA",
    "Boeing",
    "Chevron",
    "Coca-Cola",
    "JPMorgan Chase & Co"
  ),
  Ticker = c(
    "NVDA",
    "BA",
    "CVX",
    "KO",
    "JPM"
  ),
  Description = c(
    "NVIDIA sells products in the Compute & Networking and Graphics segments with a high focus on automation and AI integration",
    "Boeing manufactures and sells products and services in segments like Commercial Airplanes, Defence, and Space & Security",
    "Chevron operates in all operational segments that provide energy resources across the world",
    "Coca-Cola manufactures and sells non-alcoholic beverages worldwide",
    "JPMorgan is a Commercial, Investment, and Asset & Wealth Management bank with operations around the world"
  ),
  Industry = c(
    "Technology industry",
    "Industrial industry",
    "Energy industry",
    "Beverages industry",
    "Finance services industry"
  ),
  Shares = c(
    "200",
    "200",
    "200",
    "200",
    "200"
  ),
  Weigth = c(
    "20%",
    "20%",
    "20%",
    "20%",
    "20%"
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Portfolio Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  ) %>% 
  gt::tab_source_note(source_note = "(Yahoo Finance, n.d.)")
```

## The Portfolio's Efficient Frontier

The efficient frontier is the set of portfolios plotted against their levels of risk and expected return [@Ganti_2025]. Therefore, @Ganti_2025 suggests that a portfolio is efficient if it provides the highest return for a given level of risk. However, to find the optimal portfolio, we need to identify the line on, or above, the efficient frontier with the highest Sharpe Ratio (slope)[@Vipond_2020].

The Sharpe Ratio equation:

$$
\text{Sharpe Ratio} = \frac{E(R_P) - r_f}{\sigma(R_P)}
$$

Where:

$r_f =$ The 10-year U.S. Treasury Bond [@FRED_2026]

Currently, the portfolio's Sharpe Ratio is 0.6812, while the optimal portfolio's Sharpe Ratio is 1.0872.

As mentioned before, the goal is to achieve the most optimal portfolio by choosing one of the assessment equities, or not choosing any one of them.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyquant)
library(lubridate)
library(rvest)
library(ggrepel)
library(TTR)
library(plotly)
library(xts)
library(zoo)
library(slider)
library(jsonlite)
library(tidytext)
library(stringr)
library(broom)
library(tidymodels)
library(rsample)
library(recipes)
library(parsnip)
library(yardstick)
library(rstanarm)
library(ggnewscale)
```

```{r}
portfolio_stocks <- c("NVDA", "BA", "CVX", "KO", "JPM")


portfolio <- tidyquant::tq_get(
  portfolio_stocks,
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly # Converting to weekly data because I don't want the data to be too noisy but I need large enough sample size for ML
  ) %>% 
  dplyr::mutate(return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na()


portfolio_wide <- portfolio %>% 
  dplyr::select(date, symbol, return_log) %>% 
  tidyr::pivot_wider(names_from = symbol, values_from = return_log) 

portfolio_wide <- portfolio_wide %>% 
  dplyr::mutate(Portfolio_Return = 0.2 * portfolio_wide$NVDA + 0.2 * portfolio_wide$BA + 0.2 * portfolio_wide$CVX + 0.2 * portfolio_wide$KO + 0.2 * portfolio_wide$JPM)


returns <- portfolio_wide %>% 
  dplyr::select(-date, - Portfolio_Return)

ann_returns <- colMeans(returns) * 52
ann_cov <- stats::cov(returns, use = "pairwise.complete.obs") * 52


# Static webscraping for the Risk Free Rate
url <- "https://fred.stlouisfed.org/series/DGS10"

rf_rate <- rvest::read_html(url) %>% 
  rvest::html_element(".series-meta-observation-value") %>% # I'm using the . at the beginning b/c it's a class
  rvest::html_text(trim = TRUE) %>% 
  as.numeric()

rf_rate <- rf_rate / 100

# Grabbing the Risk Free Rate Using FRED API
r_f <- tidyquant::tq_get(
  "DGS10",
  get = "economic.data",
  from = Sys.Date() - 7,
  to = Sys.Date()
) %>% 
  dplyr::arrange(date)

r_f <- r_f[1, 3] %>% dplyr::pull(price)
r_f <- r_f / 100 # I want the Risk Free to match with the rest of the return values
```

```{r}
# Simulating different portfolio possibilities for the Efficient Frontier

set.seed(123) # For replicability

simulation <- replicate(5000, {
  rnd_vals <- runif(5) # Randomizing 5 numbers
  weights <- rnd_vals / sum(rnd_vals) # Creating weights that sum up to 1
  
  port_returns <- sum(weights * ann_returns)
  port_sd <- sqrt(sum(weights * (ann_cov %*% weights))) # The %*% is for multiplying matrecies together
  
  c(
    NVDA = weights[1],
    BA = weights[2],
    CVX = weights[3],
    KO = weights[4],
    JPM = weights[5],
    return = port_returns,
    sd = as.numeric(port_sd) # as.numeric is needed b/c we used a matrix multiplication
  )
})

simulation <- as.data.frame(t(simulation)) # The t transposes, or flips, the rows and columns


simulation <- simulation %>% 
  dplyr::mutate(sharpe_ratio = (return - r_f) / sd) # Calculating the sharpe ratio to every portfolio simulation

best_sharpe_ratio <- simulation %>% 
  dplyr::slice_max(order_by = sharpe_ratio, n = 1)

mvp <- simulation %>% 
  dplyr::slice_min(order_by = sd, n = 1) # Keeping the highest values of SD in every equity


portfolio_exp_return <- mean(portfolio_wide$Portfolio_Return) * 52
portfolio_exp_sd <- sd(portfolio_wide$Portfolio_Return) * sqrt(52)
sharpe_ratio <- (portfolio_exp_return - r_f) / portfolio_exp_sd


optimal_table <- dplyr::tibble(
  sd = seq(from = 0, to = max(simulation$sd), length.out = 100),
  return = best_sharpe_ratio$sharpe_ratio * sd + r_f
)

portfolio_table <- dplyr::tibble(
  sd = portfolio_exp_sd,
  return = portfolio_exp_return
)


simulation %>% ggplot2::ggplot(ggplot2::aes(x = sd, y = return, color = return)) +
  ggplot2::geom_point(alpha = 0.3) + 
  ggplot2::scale_color_viridis_c() +
  ggplot2::geom_point(
    data = best_sharpe_ratio,
    ggplot2::aes(x = sd, y = return),
    color = "orange",
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = best_sharpe_ratio,
    ggplot2::aes(x = sd, y = return, label = "Optimal Portfolio"),
    color = "black",
    segment.color = "black",
    min.segment.length = 0,
    nudge_x = -0.025,
    nudge_y = 0.02
  ) +
  ggplot2::geom_line(
    data = optimal_table,
    ggplot2::aes(x = sd, y = return),
    color = "orange"
  ) +
  ggplot2::geom_point(
    data = portfolio_table,
    ggplot2::aes(x = portfolio_exp_sd, y = portfolio_exp_return),
    color = "palevioletred1",
    size = 2
  ) + 
  ggrepel::geom_text_repel(
    data = portfolio_table,
    ggplot2::aes(x = portfolio_exp_sd, y = portfolio_exp_return, label = "Current Portfolio"),
    color = "black",
    segment.color = "black",
    min.segment.length = 0,
    nudge_x = 0.025,
    nudge_y = 0.025
  ) +
  ggplot2::labs(
    title = "The Efficient Frontier",
    x = "Annualized Standard Deviation (Risk)",
    y = "Annualized Expected Log Return",
    color = "Log Return"
  ) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5))
```

# The List of Potential Equities

The following table presents the list of five assessment equities:

```{r}
dplyr::tibble(
  Equity = c(
    "Tesla",
    "Enbridge",
    "Amazon",
    "Lockheed Martin Corp",
    "BWX Technologies Inc."
  ),
  Ticker = c(
    "TSLA",
    "ENB",
    "AMZN",
    "LMT",
    "BWXT"
  ),
  Description = c(
    "Tesla manufactures and sells electric vehicles, along with energy generation and storage systems, worldwide",
    "Enbridge is a Canadian company that provides energy resources to both the US and Canada to power their operations",
    "Amazon is an online retail store with some physical stores worldwide, and an expansion into database services",
    "Lockheed Martin is an aerospace and defence company that manufactures and sells weapons and vehicles worldwide",
    "BWX manufactures and sells nuclear components worldwide, both to governments and commercial operations"
  ),
  Industry = c(
    "Technology & Automotive industry",
    "Energy industry",
    "Technology & Consumer industry",
    "Defense industry",
    "Nuclear & Defence industry"
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Potential Equities Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  ) %>% 
  gt::tab_source_note(source_note = "(Yahoo Finance, n.d.)")
```

# Analysis Components

## Machine Learning Forecasting

Three models were tested, including Linear Regression, Random Forest, and Bayesian. However, the Linear Regression and Random Forest models presented a very low $R^2$, the explanatory power, which was expected as the efficient market theory should hold.

As a result, the Bayesian model was selected because it provides a prediction interval rather than a single-point estimate, which is more appropriate for decision-making under uncertainty. To better visualize the Bayesian model's predicted values, the middle value of the interval is computed and displayed as a red line.

### Technical Predictors

To fairly assess each equity, the following predictors are used to estimate next week's log returns.

```{r}
dplyr::tibble(
  Predictor = c(
    "Historical Log Returns",
    "MACD Histogram",
    "MFI",
    "Rolling SD (12)",
    "S&P 500",
    "VIX"
  ),
  Classification = c(
    "Technicals",
    "Technicals",
    "Technicals",
    "Technicals",
    "Market",
    "Market"
  ),
  Description = c(
    "Captures short-term momentum and trend of the time series",
    "The MACD histogram helps identify and measure trends in price and momentum, and to identify entry and exit points (Dolan, 2025). The histogram reflect differences between short-term and long-term moving averages and highlights increasing and decreasing trends by using darker colours (Dolan, 2025).",
    "Analyzes price movements and trading volumes to measure the flow of money into and out of an asset (Gratton, 2025).",
    "Using a moving standard deviation of the previous 12 weeks helps smooth out random spikes and capture short-term market transitions.",
    'Using "^GSPC" rather than "SPY" provides a direct comparison to the market, allowing capture of overall market movements and isolation of stock-specific returns.',
    "The Volatility Index, or the Fear Index, measures expected market volatility and is used to evaluate market sentiment and perceived risk (Kuepper, 2025)."
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Predictors Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  )
```

## Financial News Sentiment

Financial news sentiment measures the tone of financial news. Positive sentiment may signal investor optimism and future price increases, while negative sentiment signals uncertainty and a future price decrease. This sentiment analysis uses three specific lexicons for each equity:

1.  **Loughran-McDonald:** According to @R_Project, the Loughran lexicon is created to interpret financial documents by labelling words into the following categories: negative, positive, litigious, uncertainty, constraining, and superfluous.
2.  **AFINN:** According to @Ketchley_2023, AFINN captures valence, arousal, and dominance while providing a score to each word in accordance with those categories. The focus of this AFINN analysis is to capture the valence score for each word.
3.  **NRC:** @Canada_2019 suggests that NRC is a collection of seven emotional word lexicons, designed to capture the emotional tone of a text by categorizing words into categories such as positive, joy, trust, anticipation, fear, sadness, anger, disgust, negative, and surprise.

To arrive at a unified sentiment score, it computes the weighted average of the sentiment scores, assigning Loughran 85%, AFINN 10%, and NRC 5%. Additionally, the financial data for the sentiment analysis uses @YahooFinance articles around each equity, as well as macro-level economic news around the world.

$$
\text{Sentiment} = \text{Loughran} \cdot 0.85 + \text{AFINN} \cdot 0.1 + \text{NRC} \cdot 0.05
$$

```{r}
options(mc.cores = parallel::detectCores()) # Utilizing all the cores

# Dynamic Webscraping
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "C:/Program Files/nodejs", sep=";"))

# TSLA News
#out_tsla <- system("node ./scripts/TSLA/news_tsla.js", intern = TRUE)
txt_tsla <- paste(readLines("./scripts/TSLA/articles_tsla.json", warn = FALSE), collapse = "\n")

# ENB News
#out_enb <- system("node ./scripts/ENB/news_enb.js", intern = TRUE)
txt_enb <- paste(readLines("./scripts/ENB/articles_enb.json", warn = FALSE), collapse = "\n")

# AMZN News
#out_amzn <- system("node ./scripts/AMZN/news_amzn.js", intern = TRUE)
txt_amzn <- paste(readLines("./scripts/AMZN/articles_amzn.json", warn = FALSE), collapse = "\n")

# LMT News
#out_lmt <- system("node ./scripts/LMT/news_lmt.js", intern = TRUE)
txt_lmt <- paste(readLines("./scripts/LMT/articles_lmt.json", warn = FALSE), collapse = "\n")

# BWXT News
#out_bwxt <- system("node ./scripts/BWXT/news_bwxt.js", intern = TRUE)
txt_bwxt <- paste(readLines("./scripts/BWXT/articles_bwxt.json", warn = FALSE), collapse = "\n")

# Market News
#out_market <- system("node ./scripts/Market/news_market.js", intern = TRUE)
txt_market <- paste(readLines("./scripts/Market/articles_market.json", warn = FALSE), collapse = "\n")
```

```{r}
# Assigning scraped text from the saved JSON file
news_tsla <- jsonlite::fromJSON(txt_tsla)
news_enb <- jsonlite::fromJSON(txt_enb)
news_amzn <- jsonlite::fromJSON(txt_amzn)
news_lmt <- jsonlite::fromJSON(txt_lmt)
news_bwxt <- jsonlite::fromJSON(txt_bwxt)
news_market <- jsonlite::fromJSON(txt_market)

# Sentiments
loughran <- tidytext::get_sentiments("loughran")
afinn <- tidytext::get_sentiments("afinn")
nrc <- tidytext::get_sentiments("nrc")
```

```{r}
# Cleaning function
clean_text <- function(data){
  x <- data %>% 
    dplyr::mutate(
      text = tolower(text), # Converting everything to lowercase
      text = gsub(pattern = "'s", replacement = "", text), # Removing 's
      text = gsub(pattern = "[^a-z\\s]", replacement = " ", text), # Removing everything that's not letters: a-z ; spaces: \\s
      text = stringr::str_squish(text) # Removing leading and trailing spaces
      )
  
  return(x)
}

# Text removal function
text_rm <- function(data, remove){
  x <- data %>% 
    dplyr::mutate(
      text = gsub(pattern = remove, replacement = "", text, ignore.case = TRUE) # A substitution method
    )
  
  return(x)
}
```

```{r}
# TSLA
remove_tsla <- c(
  "tesla", "tsla", "spacex", "ubtech", "rivian", "company", "companies", "firm", "business", "stock", "stocks", "market", "markets", "investor", "investors", "trade", "trading", "shares", "price", "news", "said", "says", "report", "reports", "according", "reuters", "briefing", "morning", "update", "march", "april", "year", "today", "week", "previously", "earlier", "latest", "trump", "president", "war", "tariffs", "policy", "china", "iran", "global", "expected", "growth", "data", "term", "billion", "million", "energy", "nasdaq", "thursday", "including", "article", "wall", "key", "major", "appeared", "analyst", "guruocus", "xai", "tbpn", "gurufocus", "free", "warning", "community", "change", "deployments", "billions", "millions"
  ) # After inspecting the URL link I removed common finance words that could appear in any article and redundant words

clean_tsla <- news_tsla %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_tsla
  )

# ENB
remove_enb <- c(
  "free", "community", "news", "said", "report", "reports", "according", "reuters", "article", "stock", "stocks", "buy", "hold", "investor", "investors", "canadian", "tfsa", "rrsp", "top", "best", "why", "keep", "eye", "built", "right", "now", "including", "based", "around", "well", "still", "also", "analysis", "story", "stock", "stocks", "share", "price", "current", "investors", "company", "companies", "market", "enbridge", "enb", "tsx", "fool", "motley", "term", "time", "past", "list","business", "assets", "portfolio", "canada", "billions", "billion", "millions", "zacks", "makes", "nyse", "focused", "projects", "products", "enterprise", "stay", "model", "investment", "increased", "worth", "balance", "potential", "ability", "epd"
  )

clean_enb <- news_enb %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_enb
  )

# AMZN
remove_amzn <- c(
  "amazon", "amzn", "apple", "article", "analysis", "story", "stock", "stocks", "share", "price", "current", "investors", "company", "companies", "market", "narrative", "narratives", "based", "time", "analysis", "article", "story", "wall", "simply", "free", "top", "community", "globalstar", "term", "business", "analyst", "alternatively", "billions", "millions", "fair", "services", "buy", "potential", "valuation", "capital", "stake", "shares", "return", "retail", "margins", "financial", "deal", "data", "control", "support", "report", "past", "news", "nasdaq", "million", "dcf", "existing", "iphone", "satellite", "aws", "cloud", "worth", "ratio", "position", "future", "industry", "emergency", "recent", "quarter", "kuper", "portfolio", "forward", "head", "compared", "advertising", "moves", "billionaire", "businesses", "commentary", "directly", "estimate", "driven", "note", "infrastructure", "kuiper", "sits", "margins", "focused", "key", "discover", "platform", "reported", "features", "include", "sell", "investment", "tech", "power", "partnerships"
  )

remove_bigram_amzn <- c(
  "simply wall", "wall st", "data note", "cash flows", "current share", "share price", "nasdaq amzn", "million shares", "project kuiper", "historical data", "fundamental data", "financial situation", "financial advice", "team simplywallst", "research report", "qualitative material", "provide commentary", "email editorial", "editorial team", "stay updated", "sits alongside", "news stories", "unbiased methodology", "fair ratio", "multiline retail", "million individual", "past week", "nasdaqs aapl", "nasdaqs aapl", "nasdaqgs aapl", "iphone models", "assumes fair", "gains control", "low earth", "earth orbit"
  )

clean_amzn <- news_amzn %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::mutate(
    word = dplyr::case_when(
      word == "flows" ~ "flow",
      word == "risks" ~ "risk",
      word == "profit" ~ "earnings",
      TRUE ~ word
      )
    ) %>% 
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_amzn
  )

# LMT
remove_lmt <- c(
  "free", "community", "news", "said", "report", "reports", "according", "reuters", "article", "stock", "stocks", "buy", "hold", "investor", "investors", "canadian", "tfsa", "rrsp", "top", "best", "why", "keep", "eye", "built", "right", "now", "including", "based", "around", "well", "still", "also", "analysis", "story", "stock", "stocks", "share", "price", "current", "investors", "company", "companies", "market", "billions", "billion", "millions", "martin", "lockheed", "lmt", "trump", "nasa", "nyse", "artemis", "orion", "redwire", "production", "program", "programs", "system", "systems", "support", "term", "narrative", "increase", "space", "launch", "lunar", "zacks", "wall", "trillion", "simply", "moon", "president", "analysts", "crewed", "march", "center", "rtx", "worth", "past", "future", "operations", "data", "research", "key", "efforts", "expanding", "infrastructure", "financial", "human", "hardware", "spacecraft", "navigation", "discretionary", "critical", "exposure", "momentum", "airbus", "building", "businesses", "camera", "corporation", "capacity", "white", "portfolio", "noc", "iran", "include", "department", "directly", "house", "time", "successful", "world", "sun", "shares", "return", "rate", "yield", "sensors", "manufacturing", "integration", "imaging", "facility", "exploration", "friday", "government", "deep", "domestic"
  )

clean_lmt <- news_lmt %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::mutate(
    word = dplyr::case_when(
      word == "investments" ~ "investment",
      TRUE ~ word
      )
    ) %>%
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_lmt
  )

# BWXT
remove_bwxt <- c(
  "free", "community", "news", "said", "report", "reports", "according", "reuters", "article", "stock", "stocks", "buy", "hold", "investor", "investors", "canadian", "tfsa", "rrsp", "top", "best", "why", "keep", "eye", "built", "right", "now", "including", "based", "around", "well", "still", "also", "analysis", "story", "stock", "stocks", "share", "price", "current", "investors", "company", "companies", "market", "billions", "millions", "analysts", "bwx", "bwxt", "fair", "million", "zacks", "billion", "flr", "wall", "earth", "guidance", "commercial", "term", "simply", "clean", "days", "narrative", "projects", "business", "components", "services", "financial", "valuation", "estimate", "systems", "communication", "supply", "chain", "government", "time", "data", "contract", "publisher", "rank", "model", "access", "expected", "advanced", "results", "operations", "materials", "realoys", "oilprice", "technologies", "realaloys", "fluor", "contracts", "shares", "technology", "information", "project", "programs", "facility", "potential", "narratives", "global", "future", "analyst", "attention", "smr", "past", "program", "reactor", "propulsion", "visteon", "scale", "ceg", "ratio", "eps", "platforms", "facilities", "realaloys", "ceo", "capital", "constellation", "worth", "recent", "realloys", "consensus", "margins", "position", "markets"
  )

clean_bwxt <- news_bwxt %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>%
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::mutate(
    word = dplyr::case_when(
      word == "growing" ~ "growth",
      word == "contracts" ~ "contract",
      word == "narratives" ~ "narrative",
      word == "programs" ~ "program",
      TRUE ~ word
      )
    ) %>% 
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_bwxt
  )

# Market
remove_market <- c(
  "free", "community", "yahoo", "read", "click", "follow", "news", "stock", "stocks", "market", "markets", "price", "prices", "finance", "financial", "business", "company", "companies", "president", "age", "social", "security", "retirement", "whoop", "trump", "iran", "global", "million", "street", "month", "wrote", "world", "quarter", "musk", "billion", "ceo", "analyst", "spacex", "wall", "tesla", "start", "reporter", "middle", "hill", "wednesday", "america", "analysis", "americans", "said", "says", "told", "added", "noted", "according", "year", "week", "day", "time", "today", "last", "new", "data", "report", "note", "expected", "estimate", "forecast", "country", "countries", "global", "economy", "economic", "hamza", "ines", "ferre", "nike", "pvt", "strait", "hormuz", "east", "major", "recent", "public", "moving", "events", "biggest", "impact", "top", "cut", "hit", "rising", "ahmed", "depth", "benefit", "concerns", "gulf", "march", "tuesday", "delta", "airlines", "infrastructure", "product", "monthly", "months", "analysts", "aren", "senior", "expectations", "investors", "health", "weeks", "space", "investing", "round", "ipo"
  )

clean_market <- news_market %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>%
  dplyr::mutate(
    word = dplyr::case_when(
      word == "growing" ~ "growth",
      word == "contracts" ~ "contract",
      word == "narratives" ~ "narrative",
      word == "programs" ~ "program",
      word == "money" ~ "capital",
      word == "tariff" ~ "tariffs",
      TRUE ~ word
      )
    ) %>%
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_market
  )
```

```{r, include=FALSE}
clean_market %>%
  dplyr::count(word, sort = TRUE) %>% 
  dplyr::top_n(20) %>% 
  ggplot2::ggplot(ggplot2::aes(x = stats::reorder(word, n), y = n, fill = word)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::coord_flip() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::ggtitle("Most Used Words")
```

```{r, warning=FALSE}
sentiment_analysis <- function(data){
  # The Loughran Sentiment
  loughran <- data %>% 
    dplyr::inner_join(loughran, by = "word") %>% 
    dplyr::rename(loughran = sentiment) %>% 
    dplyr::count(article_num, loughran) %>% 
    tidyr::pivot_wider(
      names_from = loughran,
      values_from = n,
      values_fill = 0
    ) %>% 
    dplyr::mutate(net_loughran = positive - negative)
  
  loughran_score <- sum(loughran$net_loughran) / nrow(data)
  
  # The AFINN Sentiment
  afinn <- data %>% 
    dplyr::inner_join(afinn, by = "word") %>%
    dplyr::rename(net_afinn = value)
  
  afinn_score <- sum(afinn$net_afinn) / nrow(data)
  
  # The NRC Sentiment
  nrc <- data %>% 
    dplyr::inner_join(nrc, by = "word") %>% 
    dplyr::rename(nrc = sentiment) %>% 
    dplyr::count(article_num, nrc) %>% 
    tidyr::pivot_wider(
      names_from = nrc,
      values_from = n,
      values_fill = 0
    ) %>% 
    dplyr::mutate(total = positive + joy + trust + anticipation - fear - sadness - anger - disgust - negative)
  
  nrc_score <- sum(nrc$total) / nrow(data)
  
  # Calculation
  score <- 0.85 * loughran_score + 0.1 * afinn_score + 0.05 * nrc_score

  interpretation <- dplyr::case_when(
    score >= 0.022700 ~ "Strong Positive", # Above the 3rd quartile
    score >= 0.008919 ~ "Positive", # Between the Median and 3rd qurtile
    score >= 0 ~ "Weak Positive", # Between zero and the Median
    TRUE ~ "Negative"
  )
  
  return(list(score = score, inter = interpretation))
}
```

To interpret the final sentiment score, thresholds are based on the cross-sectional distribution of the aggregate sentiment scores for the five candidate equities and the market sentiment score.

```{r, warning=FALSE}
# Using the distribution to decide on the interpretation ranges
scores <- c(
  sentiment_analysis(clean_tsla)$score,
  sentiment_analysis(clean_enb)$score,
  sentiment_analysis(clean_amzn)$score,
  sentiment_analysis(clean_lmt)$score,
  sentiment_analysis(clean_bwxt)$score,
  sentiment_analysis(clean_market)$score
)

summary_tbl <- tibble::tibble(
  Statistic = names(summary(scores)),
  Value = round(as.numeric(summary(scores)), 6)
)

summary_tbl %>%
  gt::gt() %>% 
  gt::tab_header(title = "Sentiment Score Distribution Summary") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  )
```

-   **Score** $\geq$ **3rd quartile:** Strong positive sentiment

-   **Median** $\leq$ **Score** $<$ **3rd quartile:** Positive sentiment

-   $0 \leq$ **Score** $<$ **Median:** Weak positive sentiment

-   **Score** $< 0$**:** Negative sentiment

```{r, include=FALSE}
news_amzn %>% 
  clean_text() %>% 
  tidytext::unnest_tokens(
    output = word,
    input = text,
    token = "ngrams",
    n = 2
  ) %>%
  tidyr::separate(word, c("w1", "w2"), sep = " ") %>% 
  dplyr::filter(
    !w1 %in% stop_words$word,
    !w2 %in% stop_words$word,
    !w1 %in% remove_amzn,
    !w2 %in% remove_amzn,
    nchar(w1) > 2,
    nchar(w2) > 2
    ) %>% 
  stats::na.omit() %>% 
  tidyr::unite(word, w1, w2, sep = " ") %>% 
  dplyr::filter(!word %in% remove_bigram_amzn) %>% 
  dplyr::count(word, sort = TRUE) %>% 
  dplyr::top_n(5) %>% 
  ggplot2::ggplot(ggplot2::aes(x = stats::reorder(word, n), y = n, fill = word)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::coord_flip() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::ggtitle("Most Used Words")
```

```{r}
assessment_stocks <- c("TSLA", "ENB", "AMZN", "LMT", "BWXT")
```

```{r, warning=FALSE}
# Grabing the assessment equity data
assessment <- tidyquant::tq_get(
  assessment_stocks,
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na()

assessment_wide <- assessment %>% 
  dplyr::select(date, symbol, return_log) %>% 
  tidyr::pivot_wider(names_from = symbol, values_from = return_log)


data_vix <- tidyquant::tq_get(
  "^VIX",
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(vix_return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na() %>% 
  dplyr::rename(vix_adjusted = adjusted) %>%
  dplyr::ungroup() %>% 
  dplyr::select(date, vix_adjusted, vix_return_log)

data_gspc <- tidyquant::tq_get(
  "^GSPC",
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(gspc_return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na() %>% 
  dplyr::rename(gspc_adjusted = adjusted) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(date, gspc_adjusted, gspc_return_log)
```

```{r}
# Technicals calculation function
technicals <- function(
    ticker,
    original = assessment,
    original_wide = assessment_wide,
    market_gspc = data_gspc,
    market_vix = data_vix
    ){
  
  
  data_stock <- original %>% 
    dplyr::filter(symbol == ticker) %>% 
    dplyr::ungroup() %>% 
    dplyr::arrange(date)
  
  # MACD
  table <- TTR::MACD(
    original_wide[[ticker]],
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    maType = "EMA",
    percent = FALSE
    )
  
  table <- data_stock %>% 
    dplyr::select(date, open, close, high, low, volume, return_log) %>% 
    dplyr::left_join(
      as.data.frame(table) %>%  # Converting back to a data frame table
        dplyr::transmute(date = original_wide$date, macd, signal),
      by = "date"
      ) %>% 
    dplyr::mutate(macd_hist = macd - signal)
  
  
  # MFI
  xts_tbl <- xts::xts(
    x = data_stock %>% dplyr::select(open, high, low, close, volume),
    order.by = data_stock$date
    )

  tech_mfi <- TTR::MFI(
    HLC = xts_tbl[, c("high", "low", "close")],
    volume = xts_tbl[, "volume"],
    n = 14 # Using a 2 week period
    )

  tbl_mfi <- data.frame( 
    date = zoo::index(tech_mfi),
    zoo::coredata(tech_mfi)
    )

  table <- table %>% dplyr::left_join(tbl_mfi, by = "date")
  
  
  # Rolling SD
  table <- table %>% 
    dplyr::mutate(
      sd12 = slider::slide_dbl(
        .x = return_log,
        .f = sd,
        .before = 12, # Looking 3 months back
        .after = 0,
        .complete = FALSE 
        ) * sqrt(52)
      ) # I didn't remove NAs b/c other technicals like MACD need long beg. periods, so it will remove a lot of data
  
  
  # ^VIX and ^GSPC
  table <- table %>% 
    dplyr::left_join(market_vix, by = "date") %>% 
    dplyr::left_join(market_gspc, by = "date")
  
  
  table <- table %>% 
    dplyr::mutate(
      future_return = dplyr::lead(return_log, n = 1), # Needed for forecasting
      plot_date = dplyr::lead(date, n = 1)
    )
  
  return(list(table = table))
}
```

```{r}
# Machine Learning function for predicting the log returns
model_ml <- function(ticker){
  options(mc.cores = parallel::detectCores())
  set.seed(123)
  
  table <- technicals(ticker)$table %>% 
  tidyr::drop_na()
  
  # Sampling
  split_data <- table %>%
    dplyr::select(date, plot_date, future_return, return_log, sd12, macd_hist, mfi, vix_adjusted, gspc_return_log) %>% 
    rsample::initial_time_split(prop = 0.7)
  
  training_data <- rsample::training(split_data)
  testing_data <- rsample::testing(split_data)
  
  # Features Engineering
  recipe_pipeline <- 
    recipes::recipe(future_return ~ ., data = training_data) %>% 
    recipes::step_rm(date, plot_date) %>% 
    recipes::step_normalize(all_numeric_predictors()) %>% 
    recipes::prep()
  
  training_baked <- recipes::bake(recipe_pipeline, training_data)
  testing_baked <- recipes::bake(recipe_pipeline, testing_data)
  
  
  # Modeling - Linear Regression
  model_lm <-
    parsnip::linear_reg(mode = "regression") %>% 
    parsnip::set_engine("lm") %>% 
    parsnip::fit(future_return ~ ., data = training_baked)
  
  # Predicting
  predicted_lm <- model_lm %>% 
    stats::predict(new_data = testing_baked) %>% 
    dplyr::bind_cols(future_return = testing_baked %>% dplyr::select(future_return))
  
  # Metrics
  tidy_lm <- parsnip::tidy(model_lm) %>% 
    gt::gt() %>% 
    gt::fmt_percent(columns = matches("std.error"), decimals = 2)
  
  glance_lm <- parsnip::glance(model_lm) %>% 
    gt::gt() %>% 
    gt::fmt_percent(columns = matches("r.squared|adj.r.squared"), decimals = 2)
  
  yardstick_lm <- predicted_lm %>% 
    yardstick::metrics(truth = future_return, estimate = .pred) %>% 
    dplyr::arrange(.metric) %>% 
    gt::gt() %>% 
    gt::fmt_percent(columns = matches(".estimate"), decimals = 2)
  
  
  
  # Modeling - Random Forest
  model_rf <-
    parsnip::rand_forest(
      mode = "regression",
      trees = 750, # 750 different decision trees
      mtry = 3, # number of variables to randomly select and choosing the best split out of them
      min_n = 5 # Min number of observations in a leaf or end of a tree branch
      ) %>% 
    parsnip::set_engine(
      "ranger",
      importance = "impurity" # Using reduction in variance for each split
      ) %>% 
    parsnip::fit(future_return ~ ., data = training_baked)
  
  # Predicting
  predicted_rf <- model_rf %>% 
    stats::predict(new_data = testing_baked) %>% 
    dplyr::bind_cols(future_return = testing_baked %>% dplyr::select(future_return))
  
  # Metrics
  yardstick_rf <- predicted_rf %>% 
    yardstick::metrics(truth = future_return, estimate = .pred) %>% 
    dplyr::arrange(.metric) %>% 
    gt::gt() %>% 
    gt::fmt_percent(columns = matches(".estimate"), decimals = 2)
  
  
  
  # Modeling - Bayesian Regression
  model_bayes <-
    parsnip::linear_reg(mode = "regression") %>% 
    parsnip::set_engine(
      "stan",
      prior_intercept = rstanarm::normal(0, 1), # I believe -> mean: 0; SD: 1
      prior = rstanarm::normal(0, 0.5), # I believe -> mean: 0; SD: 0.5
      iter = 4000, # Number of simulations
      seed = 123
    ) %>% 
    parsnip::fit(future_return ~ ., data = training_baked)
  
  # Predicting
  predicted_bayes <- model_bayes %>% 
    stats::predict(new_data = testing_baked, type = "pred_int") %>% 
    dplyr::bind_cols(future_return = testing_baked %>% dplyr::select(future_return))
  
  # Metrics
  results_bayes <- predicted_bayes %>% 
    dplyr::rename(
    "Lower Bound" = .pred_lower,
    "Upper Bound" = .pred_upper,
    "Actual Return" = future_return
    ) %>% 
    dplyr::slice_head(n = 5) %>% 
    gt::gt()
  
  
  
  # Forecasting next weeks return
  forecast_row <- technicals(ticker)$table %>% 
    dplyr::select(-future_return) %>% 
    dplyr::slice_tail(n = 1) # Taking the most recent date data for forecast
  
  forecast_baked <- recipes::bake(recipe_pipeline, new_data = forecast_row)
  
  # Forecasting the value
  forecast_lm <- stats::predict(model_lm, new_data = forecast_baked)
  forecast_rf <- stats::predict(model_rf, new_data = forecast_baked)
  forecast_bayes <- stats::predict(model_bayes, new_data = forecast_baked, type = "pred_int")
  
  # Calculating values for portfolio analysis
  bayes_exp_return <- ((forecast_bayes$.pred_lower + forecast_bayes$.pred_upper) / 2) # Taking the middle value (or avg.)
  bayes_exp_sd <- ((forecast_bayes$.pred_upper - forecast_bayes$.pred_lower) / 4) # Using 4 b/c of 95% confidence, so 4 SD difference
  
  # Saving values to a table
  exp_bayes_tbl <- dplyr::tibble(
    date = forecast_row$date + 7,
    pred_upper = forecast_bayes$.pred_upper,
    pred_lower = forecast_bayes$.pred_lower,
    pred_mid = bayes_exp_return, # Not annualized
    sd = bayes_exp_sd * sqrt(52),
    return = bayes_exp_return * 52 # Annualized for the portfolio
  )
  
  
  # Creating a table so I can plot the predicted with the actual
  plot_bayes_tbl <- testing_data %>% 
    dplyr::select(plot_date, future_return) %>% 
    dplyr::bind_cols(predicted_bayes %>% dplyr::select(.pred_lower, .pred_upper)) %>%
    dplyr::rename(
      pred_upper = .pred_upper,
      pred_lower = .pred_lower
    ) %>% 
    tibble::add_row(
      plot_date = max(testing_data$plot_date, na.rm = TRUE) + 7,
      pred_upper = forecast_bayes$.pred_upper,
      pred_lower = forecast_bayes$.pred_lower
    ) %>% 
    dplyr::mutate(
      pred_mid = (pred_upper + pred_lower ) / 2
    )
  
  
  
  return(list(
    results_bayes = results_bayes,
    exp_bayes_tbl = exp_bayes_tbl,
    plot_bayes_tbl = plot_bayes_tbl
    ))
}
```

```{r}
# Plotting function
plot_it <- function(ticker){
  
  table <- technicals(ticker)$table
  
  histogram <- table %>% 
    dplyr::select(date, macd_hist) %>% 
    dplyr::mutate(
      direction = macd_hist - dplyr::lag(macd_hist),
      macd_color = dplyr::case_when(
        is.na(dplyr::lag(macd_hist)) ~ "#00FF7F",
        macd_hist >= 0 & direction >= 0 ~ "#00CD66",
        macd_hist >= 0 & direction < 0 ~ "#00FF7F",
        macd_hist < 0 & direction <= 0 ~ "#CD2626",
        macd_hist < 0 & direction > 0 ~ "#FF3030"
      )
    ) %>% 
    tidyr::drop_na()
  
  data_ml <- model_ml(ticker)$plot_bayes_tbl
  
  plot_return <- table %>% 
    plotly::plot_ly(x = ~date) %>%
    plotly::add_lines(x = ~date, y = ~return_log, name = "Log Returns") %>% 
    plotly::add_lines(
      data = data_ml,x = ~plot_date, y = ~pred_upper, name = "Predicted Upper Bound", line = list(width = 0)
      ) %>% 
    plotly::add_lines(data = data_ml, x = ~plot_date, y = ~pred_lower, name = "Predicted Lower Bound", fill = "tonexty", fillcolor = "rgba(0, 100, 200, 0.2)", line = list(width = 0)) %>% 
    plotly::add_lines(data = data_ml, x = ~plot_date, y = ~pred_mid, name = "Predicted Middle") %>% 
    plotly::layout(
      title = list(text = paste(ticker, "Time Series by Week")),
      xaxis = list(title = "Date"),
      yaxis = list(title = "Log Return", titlefont = list(size = 12))
      )

  plot_macd <- table %>%
    plotly::plot_ly(x = ~date) %>% 
    plotly::add_lines(x = ~date, y = ~macd, name = "MACD") %>% 
    plotly::add_lines(x = ~date, y = ~signal, name = "Signal") %>% 
    plotly::add_bars(data = histogram, x = ~date, y = ~macd_hist, name = "MACD Histogram", marker = list(color = ~macd_color))
  
  plot_sd12 <- table %>%
    plotly::plot_ly(
      x = ~date,
      y = ~sd12,
      type = "scatter",
      mode = "lines"
      ) %>%  
    plotly::style(name = "Rolling SD (12)")
  
  plot_mfi <- table %>% 
    plotly::plot_ly(
      x = ~date,
      y = ~mfi,
      type = "scatter",
      mode = "lines"
    ) %>%  
    plotly::style(name = "MFI")
  
  plot_mkt <- table %>% 
    plotly::plot_ly(x = ~date) %>%
    plotly::add_lines(x = ~date, y = ~return_log, name = "Log Returns") %>% 
    plotly::add_lines(x = ~date, y = ~gspc_return_log, name = "S&P 500") %>% 
    plotly::add_lines(
      data = data_ml,x = ~plot_date, y = ~pred_upper, name = "Predicted Upper Bound", line = list(width = 0)
      ) %>% 
    plotly::add_lines(data = data_ml, x = ~plot_date, y = ~pred_lower, name = "Predicted Lower Bound", fill = "tonexty", fillcolor = "rgba(0, 100, 200, 0.2)", line = list(width = 0)) %>% 
    plotly::add_lines(data = data_ml, x = ~plot_date, y = ~pred_mid, name = "Predicted Middle", line = list(color = "red"))
  
  plot_vix <- table %>% 
    plotly::plot_ly(
      x = ~date,
      y = ~vix_adjusted,
      type = "scatter",
      mode = "lines"
    ) %>%  
    plotly::style(name = "VIX")
  
  
  plot_all_tech <- plotly::subplot(
    plot_return,
    plot_macd,
    plot_sd12,
    plot_mfi,
    nrows = 4,
    heights = c(0.30, 0.30, 0.20, 0.20), # Top plot get 30% of height, 2nd 30%, 3rd 20%, and bottom plot gets 20%
    shareX = TRUE,
    titleX = TRUE
    ) %>% 
    plotly::layout(
      title = list(text = paste(ticker, "Time Series by Week")),
      xaxis = list(title = "Date"),
      yaxis = list(title = "Log Return", titlefont = list(size = 12)),
      yaxis2 = list(title = "EMA", titlefont = list(size = 10)),
      yaxis3 = list(title = "Standard Deviation", titlefont = list(size = 10)),
      yaxis4 = list(title = "Momentum", titlefont = list(size = 10))
      )
  
  
  plot_all_mkt <- plotly::subplot(
    plot_mkt,
    plot_vix,
    nrows = 2,
    heights = c(0.55, 0.45),
    shareX = TRUE,
    titleX = TRUE
    ) %>% 
    plotly::layout(
      title = list(text = paste(ticker, "Time Series by Week")),
      xaxis = list(title = "Date"),
      yaxis = list(title = "Log Return", titlefont = list(size = 12)),
      yaxis2 = list(title = "Value", titlefont = list(size = 10))
      )
  
  
  return(list(
    plot_return = plot_return,
    plot_stock = plot_all_tech,
    plot_market = plot_all_mkt
    ))
}
```

# The Analysis

```{r}
# Portfolio Correlation Function
portfolio_cor <- function(ticker, port_long = portfolio, port_wide = portfolio_wide){
  new_stock_hist <- technicals(ticker)$table %>% 
    dplyr::select(date, return_log) %>% 
    dplyr::rename(!!sym(ticker) := return_log) # Converting the ticket string to a symbol and using walrus for assignment
  
  port_hist <- port_wide %>% 
    dplyr::select(date, NVDA, BA, CVX, KO, JPM) %>% 
    dplyr::left_join(new_stock_hist, by = "date") %>% 
    dplyr::select(-date)
  
  port_cor <- stats::cor(port_hist, use = "complete.obs") %>% 
    corrplot::corrplot(
    method = "color",
    type = "upper",
    tl.col = "black",
    tl.srt = 45,
    order= "hclust",
    addCoef.col = "black"
    )
}
```

## Tesla

Looking at Tesla, the Bayesian model predicts a weekly return of -3.97% with a wide prediction interval of \[-21.97%, 13.78%\], indicating substantial uncertainty in the forecast. The wide interval is expected, as Tesla is well known for its high volatility, suggesting it may add significant risk to the portfolio.

The sentiment analysis score of 0.00188 indicates a weak positive sentiment around Tesla's financial news. The difference between the negative forecast and the weak positive sentiment suggests that the forecast is capturing market dynamics, such as momentum and investors' fear, that are not captured in the sentiment analysis.

Given the combination of negative expected return and weak positive sentiment, Tesla is unlikely to improve the portfolio's risk-adjusted performance.

```{r}
plot_tesla <- plot_it("TSLA")
```

```{r}
plot_tesla$plot_return
```

The correlation matrix below shows the relationships between the portfolio assets and the assessment equities. Lower correlation between assets provides diversification benefits, reducing overall portfolio risk.

Notably, Tesla has a relatively low correlation with several existing portfolio equities, particularly CVX(0.08) and KO (0.09), suggesting potential diversification benefits.

```{r}
portfolio_cor("TSLA")
```

```{r, warning=FALSE}
sentiment_tsla <- sentiment_analysis(clean_tsla)
```

```{r, eval=FALSE}
sentiment_tsla
```

## Enbridge

Looking at Enbridge, the Bayesian model predicts a large weekly return of 0.68% with a prediction interval of \[-4.80%, 6.17%\], indicating a moderate uncertainty in the forecast. Compared to other equities, Enbridge has the narrowest interval, suggesting stable, predictable returns.

The sentiment analysis score of 0.03398 indicates a strong positive sentiment around Enbridge's financial news. The strong positive forecast and the strong positive sentiment suggest that there is a huge upside potential for Enbridge.

```{r}
plot_enb <- plot_it("ENB")
```

```{r}
plot_enb$plot_return
```

Looking at the correlation matrix below, Enbridge has relatively low correlations with NVDA (0.13) and KO (0.25), indicating strong diversification benefits. However, ENB shows a high correlation with CVX (0.52), reflecting their shared exposure to the energy industry.

```{r}
portfolio_cor("ENB")
```

Overall, ENB provides a favourable combination of positive expected return and diversification benefits.

```{r, warning=FALSE}
sentiment_enb <- sentiment_analysis(clean_enb)
```

```{r, eval=FALSE}
sentiment_enb
```

## Amazon

Looking at Amazon, the Bayesian model predicts a weekly return of -0.31% with a prediction interval of \[-9.68%, 9.06%\], indicating a high uncertainty in the forecast.

The sentiment analysis score of 0.01286 indicates a positive sentiment around Amazon's financial news. The difference between the slight negative forecast and the positive sentiment suggests that the forecast is capturing market dynamics, such as momentum and investors' fear, that are not captured in the sentiment analysis.

```{r}
plot_amzn <- plot_it("AMZN")
```

```{r}
plot_amzn$plot_return
```

Looking at the correlation matrix below, Amazon has relatively low correlations with KO (0.05) and CVX (0.07), indicating strong diversification benefits. However, AMZN shows a high correlation with NVDA (0.57), reflecting their shared exposure to the technology industry.

```{r}
portfolio_cor("AMZN")
```

```{r, warning=FALSE}
sentiment_amzn <- sentiment_analysis(clean_amzn)
```

```{r, eval=FALSE}
sentiment_amzn
```

## Lockheed Martin

For Lockheed Martin, the Bayesian model predicts a weekly return of 0.87% with a prediction interval of \[-5.56%, 7.31%\], indicating moderate uncertainty in the forecast. While the expected return is positive, LMT still shows some short-term performance variability.

The sentiment analysis score of 0.00498 indicates a weak positive sentiment around Lockheed Martin's financial news, aligning with the model's positive forecast and suggesting a mildly favourable outlook.

```{r}
plot_lmt <- plot_it("LMT")
```

```{r}
plot_lmt$plot_return
```

Looking at the correlation matrix below, Lockheed Martin has relatively low correlations with the portfolio, particularly with NVDA (0.08) and BA (0.09); on the other hand, LMT's highest correlation with the portfolio's assets is CVX (0.3), indicating an extremely strong diversification benefits.

```{r}
portfolio_cor("LMT")
```

Overall, while LMT does not provide the strongest return, its low correlation with the portfolio makes it a valuable diversification equity. As a result, LMT has the potential to improve the portfolio's risk-adjusted return by reducing the portfolio's volatility.

```{r, warning=FALSE}
sentiment_lmt <- sentiment_analysis(clean_lmt)
```

```{r, eval=FALSE}
sentiment_lmt
```

## BWX Technologies

For BWX Technologies, the Bayesian model predicts a weekly return of 0.43% with a prediction interval of \[-6.81%, 7.68%\], indicating a moderate uncertainty in the forecast.

Although BWX's forecasted return is low and its volatility is moderate, its sentiment analysis score of 0.02598 indicates strong positive sentiment around BWX's financial news. The difference between the slight positive forecast and the strong positive sentiment suggests that the forecast is capturing market dynamics, such as momentum and investors' fear, that are not captured in the sentiment analysis.

```{r}
plot_bwxt <- plot_it("BWXT")
```

```{r}
plot_bwxt$plot_return
```

BWX Technologies possesses a relatively high correlation with the portfolio when compared to the other assessment equities, but yet a still fairly low correlation. BWXT has relatively low correlations with KO (0.12) and BA (0.23), indicating moderate diversification benefits.

```{r}
portfolio_cor("BWXT")
```

Ultimately, even though BWX's forecasted return is low, it has a strong news sentiment and a low correlation with the portfolio. Therefore, BWX could provide a great addition to the current portfolio.

```{r, warning=FALSE}
sentiment_bwxt <- sentiment_analysis(clean_bwxt)
```

```{r, eval=FALSE}
sentiment_bwxt
```

# Conclusion

Using the forecasted return for each assessment equity as their expected return and computing the optimal weights for the portfolio when each one of the equities is added to the portfolio, it is noticeable that Lockheed Martin provides the most optimal risk-adjusted return portfolio with only a slight more risk than ENB's portfolio but with much more return.

```{r}
optimal_port <- function(ticker, risk_free = r_f, port_long = portfolio, port_wide = portfolio_wide){
  set.seed(123)
  
  # Setting up all the exp ann. returns for the portfolio
  port_stock <- c(
    nvda = mean(portfolio_wide$NVDA, na.rm = TRUE) * 52,
    ba = mean(portfolio_wide$BA, na.rm = TRUE) * 52,
    cvx = mean(portfolio_wide$CVX, na.rm = TRUE) * 52,
    ko = mean(portfolio_wide$KO, na.rm = TRUE) * 52,
    jpm = mean(portfolio_wide$JPM, na.rm = TRUE) * 52,
    new_stock = model_ml(ticker)$exp_bayes_tbl$return # The annualized exp return
  )
  
  
  # Calculating the Covarince for the new portfolio
  new_stock_hist <- technicals(ticker)$table %>% 
    dplyr::select(date, return_log) %>% 
    dplyr::rename(new_stock = return_log)
  
  port_hist <- port_wide %>% 
    dplyr::select(date, NVDA, BA, CVX, KO, JPM) %>% 
    dplyr::left_join(new_stock_hist, by = "date") %>% 
    dplyr::select(-date)
  
  annual_cov <- stats::cov(port_hist, use = "complete.obs") * 52
  
  
  # Running simulation to find the optimal portfolio values
  sim <- replicate(5000, {
    rnd_vals <- runif(6)
    weight <- rnd_vals / sum(rnd_vals)
    
    port_rets <- sum(weight * port_stock)
    port_sd <- sqrt(sum(weight * (annual_cov %*% weight)))
    
    c(
      NVDA = weight[1],
      BA = weight[2],
      CVX = weight[3],
      KO = weight[4],
      JPM = weight[5],
      new_stock = weight[6],
      return = port_rets,
      sd = as.numeric(port_sd)
    )
  })
  
  sim <- as.data.frame(t(sim))
  
  sim <- sim %>% 
    dplyr::mutate(sharpe_ratio = (return - risk_free) / sd)
  
  opt_port <- sim %>% 
    dplyr::slice_max(order_by = sharpe_ratio, n = 1)
  
  port_tbl <- tibble::tibble(
    equity = ticker,
    return = opt_port$return,
    sd = opt_port$sd,
    sharpe_ratio = opt_port$sharpe_ratio,
    NVDA_w = opt_port$NVDA,
    BA_w = opt_port$BA,
    CVX_w = opt_port$CVX,
    KO_w = opt_port$KO,
    JPM_w = opt_port$JPM,
    new_stock_w = opt_port$new_stock
  )
  
  
  return(port_tbl)
}
```

```{r}
# Portfolio calculation function
portfolio_calc <- function(ticker, risk_free = r_f, port_long = portfolio, port_wide = portfolio_wide){
  set.seed(123)
  
  # Setting up all the exp ann. returns for the portfolio
  port_stock <- c(
    nvda = mean(portfolio_wide$NVDA, na.rm = TRUE) * 52,
    ba = mean(portfolio_wide$BA, na.rm = TRUE) * 52,
    cvx = mean(portfolio_wide$CVX, na.rm = TRUE) * 52,
    ko = mean(portfolio_wide$KO, na.rm = TRUE) * 52,
    jpm = mean(portfolio_wide$JPM, na.rm = TRUE) * 52,
    new_stock = model_ml(ticker)$exp_bayes_tbl$return # The annualized exp return
  )
  
  
  # Calculating the Covarince for the new portfolio
  new_stock_hist <- technicals(ticker)$table %>% 
    dplyr::select(date, return_log) %>% 
    dplyr::rename(new_stock = return_log)
  
  port_hist <- port_wide %>% 
    dplyr::select(date, NVDA, BA, CVX, KO, JPM) %>% 
    dplyr::left_join(new_stock_hist, by = "date") %>% 
    dplyr::select(-date)
  
  annual_cov <- stats::cov(port_hist, use = "complete.obs") * 52
  
  
  # Running simulation to find the optimal portfolio values
  sim <- replicate(5000, {
    rnd_vals <- runif(6)
    weight <- rnd_vals / sum(rnd_vals)
    
    port_rets <- sum(weight * port_stock)
    port_sd <- sqrt(sum(weight * (annual_cov %*% weight)))
    
    c(
      NVDA = weight[1],
      BA = weight[2],
      CVX = weight[3],
      KO = weight[4],
      JPM = weight[5],
      new_stock = weight[6],
      return = port_rets,
      sd = as.numeric(port_sd)
    )
  })
  
  sim <- as.data.frame(t(sim))
  
  sim <- sim %>% 
    dplyr::mutate(sharpe_ratio = (return - risk_free) / sd)
  
  opt_port <- sim %>% 
    dplyr::slice_max(order_by = sharpe_ratio, n = 1)
  
  port_tbl <- tibble::tibble(
    equity = ticker,
    return = opt_port$return,
    sd = opt_port$sd,
    sharpe_ratio = opt_port$sharpe_ratio,
    NVDA_w = opt_port$NVDA,
    BA_w = opt_port$BA,
    CVX_w = opt_port$CVX,
    KO_w = opt_port$KO,
    JPM_w = opt_port$JPM,
    new_stock_w = opt_port$new_stock
  )
  
  
  return(port_tbl)
}
```

```{r}
plot_tsla <- portfolio_calc("TSLA") %>% dplyr::rename(TSLA_w = new_stock_w)
plot_enb <- portfolio_calc("ENB") %>% dplyr::rename(ENB_w = new_stock_w)
plot_amzn <- portfolio_calc("AMZN") %>% dplyr::rename(AMZN_w = new_stock_w)
plot_lmt <- portfolio_calc("LMT") %>% dplyr::rename(LMT_w = new_stock_w)
plot_bwxt <- portfolio_calc("BWXT") %>% dplyr::rename(BWXT_w = new_stock_w)

plots_tbl <- dplyr::bind_rows(
  plot_tsla,
  plot_enb,
  plot_amzn,
  plot_lmt,
  plot_bwxt
)

tbl_w <- plots_tbl %>%
  dplyr::select(TSLA_w, ENB_w, AMZN_w, LMT_w, BWXT_w) %>% 
  tidyr::pivot_longer(
    c(TSLA_w, ENB_w, AMZN_w, LMT_w, BWXT_w),
    names_to = "names",
    values_to = "own_weight"
  ) %>% 
  tidyr::drop_na() %>% 
  dplyr::select(-names)

plots_tbl <- plots_tbl %>% 
  dplyr::select(-c(TSLA_w, ENB_w, AMZN_w, LMT_w, BWXT_w)) %>% 
  dplyr::bind_cols(tbl_w)
```

```{r}
simulation %>% ggplot2::ggplot(ggplot2::aes(x = sd, y = return, color = return)) +
  ggplot2::geom_point(alpha = 0.3) + 
  ggplot2::scale_color_viridis_c(name = "Log Return") +
  ggnewscale::new_scale_color() +
  ggplot2::geom_point(
    data = best_sharpe_ratio,
    ggplot2::aes(x = sd, y = return),
    color = "orange",
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = best_sharpe_ratio,
    ggplot2::aes(x = sd, y = return, label = "Optimal Portfolio"),
    color = "black",
    segment.color = "black",
    min.segment.length = 0,
    nudge_x = -0.025,
    nudge_y = 0.02
  ) +
  ggplot2::geom_line(
    data = optimal_table,
    ggplot2::aes(x = sd, y = return),
    color = "orange"
  ) +
  ggplot2::geom_point(
    data = portfolio_table,
    ggplot2::aes(x = portfolio_exp_sd, y = portfolio_exp_return),
    color = "palevioletred1",
    size = 2
  ) + 
  ggrepel::geom_text_repel(
    data = portfolio_table,
    ggplot2::aes(x = portfolio_exp_sd, y = portfolio_exp_return, label = "Current Portfolio"),
    color = "black",
    segment.color = "black",
    min.segment.length = 0,
    nudge_x = 0.045,
    nudge_y = 0.025
  ) +
  ggplot2::geom_point(
    data = plots_tbl,
    ggplot2::aes(x = sd, y = return, color = equity),
    size = 2
  ) + 
  ggplot2::scale_color_manual(
    name = "Equity",
    values = c(TSLA = "lightgoldenrod1", ENB = "tan3", AMZN = "cadetblue1", LMT = "slateblue1", BWXT = "deeppink3")
  ) +
  ggrepel::geom_text_repel(
    data = plot_lmt,
    ggplot2::aes(x = sd, y = return, label = "LMT Portfolio"),
    color = "black",
    segment.color = "black",
    min.segment.length = 0,
    nudge_x = -0.025,
    nudge_y = 0.025
  ) +
  ggplot2::labs(
    title = "The Efficient Frontier",
    x = "Annualized Standard Deviation (Risk)",
    y = "Annualized Expected Log Return",
    color = "Log Return"
  ) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5))
```

As depicted in the table below, Lockheed Martin's Sharpe Ratio is 1.7517.

That said, it is also important to consider financial news sentiment when making a decision. As mentioned before, LMT has a weak positive financial news sentiment, but since its Sharpe Ratio is much larger than the second largest Sharpe Ratio, I would suggest adding LMT to the current portfolio with the following weightings:

-   **LMT:** 44.45%

-   **NVDA:** 17.84%

-   **BA:** 0.18%

-   **CVX:** 2.34%

-   **KO:** 19.46%

-   **JPM:** 15.74%

```{r}
plots_tbl %>%
  dplyr::rename(
    Equity = equity,
    Return = return,
    SD = sd,
    `Sharpe Ratio` = sharpe_ratio,
    `NVDA Weight` = NVDA_w,
    `BA Weight` = BA_w,
    `CVX Weight` = CVX_w,
    `KO Weight` = KO_w,
    `JPM Weight` = JPM_w,
    `Own Weight` = own_weight
  ) %>% 
  dplyr::arrange(desc(`Sharpe Ratio`)) %>% 
  gt::gt() %>% 
  gt::fmt_number(columns = `Sharpe Ratio`, decimals = 4) %>% 
  gt::fmt_percent(
    columns = c(Return, SD, `NVDA Weight`, `BA Weight`, `CVX Weight`, `KO Weight`, `JPM Weight`, `Own Weight`),
    decimals = 2
  ) %>% 
  gt::tab_header(title = "The Optimal Portfolio Possibilities") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  )
```
