---
title: "Equity Portfolio Decision Making"
author: Daniel Rabi
date: April 9, 2026
backgroundcolor: lightblue
format: html
editor: visual
embed-resources: true
bibliography: references/references.bib
csl: references/apa.csl
output-file: index
nocite: "@YahooFinance, @Gratton_2025, @Dolan_2025, @Kuepper_2025"

echo: FALSE
message: FALSE
warning: FALSE
---

![[@Zerdzicki_2026]](images/into_picture.jpg){fig-align="center" width="600"}

Have you ever wondered how to decide whether you should add equity to your existing portfolio?

In this project, I will walk you through the decision-making process for assessing equities before adding them to an existing portfolio.

# The Portfolio

For this project's portfolio, I used a fixed, fairly diversified existing portfolio that contains the following companies:

```{r, message=FALSE, warning=FALSE}
library(gt)

dplyr::tibble(
  Equity = c(
    "NVIDIA",
    "Boeing",
    "Chevron",
    "Coca-Cola",
    "JPMorgan Chase & Co"
  ),
  Ticker = c(
    "NVDA",
    "BA",
    "CVX",
    "KO",
    "JPM"
  ),
  Description = c(
    "NVIDIA sells products in the Compute & Networking and Graphics segments with a high focus on automation and AI integration",
    "Boeing manufactures and sells products and services in segments like Commercial Airplanes, Defence, and Space & Security",
    "Chevron operates in all operational segments that provide energy resources across the world",
    "Coca-Cola manufactures and sells non-alcoholic beverages worldwide",
    "JPMorgan is a Commercial, Investment, and Asset & Wealth Management bank with operations around the world"
  ),
  Industry = c(
    "Technology industry",
    "Industrial industry",
    "Energy industry",
    "Beverages industry",
    "Finance services industry"
  ),
  Shares = c(
    "200",
    "200",
    "200",
    "200",
    "200"
  ),
  Weigth = c(
    "20%",
    "20%",
    "20%",
    "20%",
    "20%"
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Portfolio Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  ) %>% 
  gt::tab_source_note(source_note = "(Yahoo Finance, n.d.)")
```

## The Portfolio's Efficient Frontier

The efficient frontier is the set of portfolios plotted against their levels of risk and expected return [@Ganti_2025]. Therefore, @Ganti_2025 suggests that a portfolio is efficient if it provides the highest return for a given level of risk. However, to find the optimal portfolio, we need to identify the line with the highest Sharpe Ratio (slope) on the efficient frontier [@Vipond_2020].

The Sharpe Ratio equation:

$$
\text{Sharpe Ratio} = \frac{E(R_P) - r_f}{\sigma(R_P)}
$$

Where:

$r_f =$ The 10-year U.S. Treasury Bond [@FRED_2026]

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyquant)
library(lubridate)
library(rvest)
library(ggrepel)
library(TTR)
library(plotly)
library(xts)
library(zoo)
library(slider)
library(jsonlite)
library(tidytext)
library(stringr)
```

```{r}
portfolio_stocks <- c("NVDA", "BA", "CVX", "KO", "JPM")


portfolio <- tidyquant::tq_get(
  portfolio_stocks,
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly # Converting to weekly data because I don't want the data to be too noisy but I need large enough sample size for ML
  ) %>% 
  dplyr::mutate(return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na()


portfolio_wide <- portfolio %>% 
  dplyr::select(date, symbol, return_log) %>% 
  tidyr::pivot_wider(names_from = symbol, values_from = return_log) 

portfolio_wide <- portfolio_wide %>% 
  dplyr::mutate(Portfolio_Return = 0.2 * portfolio_wide$NVDA + 0.2 * portfolio_wide$BA + 0.2 * portfolio_wide$CVX + 0.2 * portfolio_wide$KO + 0.2 * portfolio_wide$JPM)


returns <- portfolio_wide %>% 
  dplyr::select(-date, - Portfolio_Return)

ann_returns <- colMeans(returns) * 52
ann_cov <- stats::cov(returns, use = "pairwise.complete.obs") * 52


# Static webscraping for the Risk Free Rate

url <- "https://fred.stlouisfed.org/series/DGS10"

rf_rate <- rvest::read_html(url) %>% 
  rvest::html_element(".series-meta-observation-value") %>% # I'm using the . at the beginning b/c it's a class
  rvest::html_text(trim = TRUE) %>% 
  as.numeric()

rf_rate <- rf_rate / 100

# Grabbing the Risk Free Rate Using FRED API

r_f <- tidyquant::tq_get(
  "DGS10",
  get = "economic.data",
  from = Sys.Date() - 7,
  to = Sys.Date()
) %>% 
  dplyr::arrange(date)

r_f <- r_f[1, 3] %>% dplyr::pull(price)
r_f <- r_f / 100 # I want the Risk Free to match with the rest of the return values
```

```{r}
# Simulating different portfolio possibilities for the Efficient Frontier

set.seed(123) # For replicability

simulation <- replicate(5000, {
  rnd_vals <- runif(5) # Randomizing 5 numbers
  weights <- rnd_vals / sum(rnd_vals) # Creating weights that sum up to 1
  
  port_returns <- sum(weights * ann_returns)
  port_sd <- sqrt(sum(weights * (ann_cov %*% weights))) # The %*% is for multiplying matrecies together
  
  c(
    NVDA = weights[1],
    BA = weights[2],
    CVX = weights[3],
    KO = weights[4],
    JPM = weights[5],
    return = port_returns,
    sd = as.numeric(port_sd) # as.numeric is needed b/c we used a matrix multiplication
  )
})

simulation <- as.data.frame(t(simulation)) # The t transposes, or flips, the rows and columns


simulation <- simulation %>% 
  dplyr::mutate(sharpe_ratio = (return - r_f) / sd) # Calculating the sharpe ratio to every portfolio simulation

best_sharpe_ratio <- simulation %>% 
  dplyr::slice_max(order_by = sharpe_ratio, n = 1)

mvp <- simulation %>% 
  dplyr::slice_min(order_by = sd, n = 1) # Keeping the highest values of SD in every equity

portfolio_exp_return <- mean(portfolio_wide$Portfolio_Return) * 52
portfolio_exp_sd <- sd(portfolio_wide$Portfolio_Return) * sqrt(52)
sharpe_ratio <- (portfolio_exp_return - r_f) / portfolio_exp_sd


optimal_table <- dplyr::tibble(
  sd = seq(from = 0, to = max(simulation$sd), length.out = 100),
  return = best_sharpe_ratio$sharpe_ratio * sd + r_f
)

portfolio_table <- dplyr::tibble(
  sd = portfolio_exp_sd,
  return = portfolio_exp_return
)


simulation %>% ggplot2::ggplot(ggplot2::aes(x = sd, y = return, color = return)) +
  ggplot2::geom_point(alpha = 0.3) + 
  ggplot2::scale_color_viridis_c() +
  ggplot2::geom_point(
    data = best_sharpe_ratio,
    ggplot2::aes(x = sd, y = return),
    color = "orange",
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = best_sharpe_ratio,
    ggplot2::aes(x = sd, y = return, label = "Optimal Portfolio"),
    color = "black"
  ) +
  ggplot2::geom_line(
    data = optimal_table,
    ggplot2::aes(x = sd, y = return),
    color = "orange"
  ) +
  ggplot2::geom_point(
    data = portfolio_table,
    ggplot2::aes(x = portfolio_exp_sd, y = portfolio_exp_return),
    color = "palevioletred1",
    size = 2
  ) + 
  ggrepel::geom_text_repel(
    data = portfolio_table,
    ggplot2::aes(x = portfolio_exp_sd, y = portfolio_exp_return, label = "Current Portfolio"),
    color = "black"
  ) +
  ggplot2::labs(
    title = "The Efficient Frontier",
    x = "Annualized Standard Deviation (Risk)",
    y = "Annualized Expected Log Return",
    color = "Log Return"
  ) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5))
```

# The List of Potential Equities

The five equities that I have decided to assess are in the following table:

```{r}
dplyr::tibble(
  Equity = c(
    "Tesla",
    "Enbridge",
    "Amazon",
    "Lockheed Martin Corp",
    "BWX Technologies Inc."
  ),
  Ticker = c(
    "TSLA",
    "ENB",
    "AMZN",
    "LMT",
    "BWXT"
  ),
  Description = c(
    "Tesla manufactures and sells electric vehicles, along with energy generation and storage systems, worldwide",
    "Enbridge is a Canadian company that provides energy resources to both the US and Canada to power their operations",
    "Amazon is an online retail store with some physical stores worldwide, and an expansion into database services",
    "Lockheed Martin is an aerospace and defence company that manufactures and sells weapons and vehicles worldwide",
    "BWX manufactures and sells nuclear components worldwide, both to governments and commercial operations"
  ),
  Industry = c(
    "Technology & Automotive industry",
    "Energy industry",
    "Technology & Consumer industry",
    "Defense industry",
    "Nuclear & Defence industry"
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Potential Equities Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  ) %>% 
  gt::tab_source_note(source_note = "(Yahoo Finance, n.d.)")
```

# Analysis Components

## Technical Predictors

To fairly assess each equity, I aim to estimate where each equity is heading one week ahead. Therefore, I use the following predictors to estimate next week's log returns for each equity.

```{r}
dplyr::tibble(
  Predictor = c(
    "Historical Log Returns",
    "MACD",
    "MFI",
    "Rolling SD (12)",
    "S&P 500",
    "VIX"
  ),
  Classification = c(
    "Technicals",
    "Technicals",
    "Technicals",
    "Technicals",
    "Market",
    "Market"
  ),
  Description = c(
    "Captures short-term momentum and trend of the time series",
    "The Moving Average Convergence Divergence helps identify and measure trends in price and momentum, and to identify entry and exit points (Dolan, 2025).",
    "Analyzes price movements and trading volumes to measure the flow of money into and out of an asset (Gratton, 2025).",
    "Using a moving standard deviation of the previous 12 weeks helps smooth out random spikes and capture short-term market transitions.",
    'Using "^GSPC" rather than "SPY" provides a direct comparison to the market, allowing capture of overall market movements and isolation of stock-specific returns.',
    "The Volatility Index, or the Fear Index, measures expected market volatility and is used to evaluate market sentiment and perceived risk (Kuepper, 2025)."
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Predictors Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  )
```

## Financial News Sentiment

Financial news sentiment measures the tone of financial news. Positive sentiment may signal investor optimism and future price increases, while negative sentiment signals uncertainty and a future price decrease. This sentiment analysis uses three specific lexicons for each equity:

1.  **Loughran-McDonald:** According to @R_Project, the Loughran lexicon is created to interpret financial documents by labelling words into the following categories: negative, positive, litigious, uncertainty, constraining, and superfluous.
2.  **AFINN:** According to @Ketchley_2023, AFINN captures valence, arousal, and dominance while providing a score to each word in accordance with those categories. The focus of this AFINN analysis is to capture the valence score for each word.
3.  **NRC:** @Canada_2019 suggests that NRC is a collection of seven emotional word lexicons, with the aim of capturing the emotional tone of a text by categorizing words into categories such as positive, joy, trust, anticipation, fear, sadness, anger, disgust, negative, and surprise.

To arrive at a unified sentiment score, I decided to take the weighted average of the sentiment scores, assigning Loughran 85%, AFINN 10%, and NRC 5%. Additionally, I used @YahooFinance articles as my source of financial news.

$$
\text{Sentiment} = \text{Loughran} \cdot 0.85 + \text{AFINN} \cdot 0.1 + \text{NRC} \cdot 0.05
$$

```{r}
options(mc.cores = parallel::detectCores()) # Utilizing all the cores

# Dynamic Webscraping
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "C:/Program Files/nodejs", sep=";"))

# TSLA News
#out_tsla <- system("node ./scripts/TSLA/news_tsla.js", intern = TRUE)
txt_tsla <- paste(readLines("./scripts/TSLA/articles_tsla.json", warn = FALSE), collapse = "\n")

# ENB News
#out_enb <- system("node ./scripts/ENB/news_enb.js", intern = TRUE)
txt_enb <- paste(readLines("./scripts/ENB/articles_enb.json", warn = FALSE), collapse = "\n")

# AMZN News
#out_amzn <- system("node ./scripts/AMZN/news_amzn.js", intern = TRUE)
txt_amzn <- paste(readLines("./scripts/AMZN/articles_amzn.json", warn = FALSE), collapse = "\n")

# LMT News
#out_lmt <- system("node ./scripts/LMT/news_lmt.js", intern = TRUE)
txt_lmt <- paste(readLines("./scripts/LMT/articles_lmt.json", warn = FALSE), collapse = "\n")

# BWXT News
#out_bwxt <- system("node ./scripts/BWXT/news_bwxt.js", intern = TRUE)
txt_bwxt <- paste(readLines("./scripts/BWXT/articles_bwxt.json", warn = FALSE), collapse = "\n")

# Market News
#out_market <- system("node ./scripts/Market/news_market.js", intern = TRUE)
txt_market <- paste(readLines("./scripts/Market/articles_market.json", warn = FALSE), collapse = "\n")
```

```{r}
# Assigning scraped text from the saved JSON file
news_tsla <- jsonlite::fromJSON(txt_tsla)
news_enb <- jsonlite::fromJSON(txt_enb)
news_amzn <- jsonlite::fromJSON(txt_amzn)
news_lmt <- jsonlite::fromJSON(txt_lmt)
news_bwxt <- jsonlite::fromJSON(txt_bwxt)
news_market <- jsonlite::fromJSON(txt_market)

# Sentiments
loughran <- tidytext::get_sentiments("loughran")
afinn <- tidytext::get_sentiments("afinn")
nrc <- tidytext::get_sentiments("nrc")
```

```{r}
# Cleaning function
clean_text <- function(data){
  x <- data %>% 
    dplyr::mutate(
      text = tolower(text), # Converting everything to lowercase
      text = gsub(pattern = "'s", replacement = "", text), # Removing 's
      text = gsub(pattern = "[^a-z\\s]", replacement = " ", text), # Removing everything that's not letters: a-z ; spaces: \\s
      text = stringr::str_squish(text) # Removing leading and trailing spaces
      )
  
  return(x)
}

# Text removal function
text_rm <- function(data, remove){
  x <- data %>% 
    dplyr::mutate(
      text = gsub(pattern = remove, replacement = "", text, ignore.case = TRUE) # A substitution method
    )
  
  return(x)
}
```

```{r}
# TSLA
remove_tsla <- c(
  "tesla", "tsla", "spacex", "ubtech", "rivian", "company", "companies", "firm", "business", "stock", "stocks", "market", "markets", "investor", "investors", "trade", "trading", "shares", "price", "news", "said", "says", "report", "reports", "according", "reuters", "briefing", "morning", "update", "march", "april", "year", "today", "week", "previously", "earlier", "latest", "trump", "president", "war", "tariffs", "policy", "china", "iran", "global", "expected", "growth", "data", "term", "billion", "million", "energy", "nasdaq", "thursday", "including", "article", "wall", "key", "major", "appeared", "analyst", "guruocus", "xai", "tbpn", "gurufocus", "free", "warning", "community", "change", "deployments", "billions", "millions"
  ) # After inspecting the URL link I removed common finance words that could appear in any article and redundant words

clean_tsla <- news_tsla %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_tsla
  )

# ENB
remove_enb <- c(
  "free", "community", "news", "said", "report", "reports", "according", "reuters", "article", "stock", "stocks", "buy", "hold", "investor", "investors", "canadian", "tfsa", "rrsp", "top", "best", "why", "keep", "eye", "built", "right", "now", "including", "based", "around", "well", "still", "also", "analysis", "story", "stock", "stocks", "share", "price", "current", "investors", "company", "companies", "market", "enbridge", "enb", "tsx", "fool", "motley", "term", "time", "past", "list","business", "assets", "portfolio", "canada", "billions", "billion", "millions", "zacks", "makes", "nyse", "focused", "projects", "products", "enterprise", "stay", "model", "investment", "increased", "worth", "balance", "potential", "ability", "epd"
  )

clean_enb <- news_enb %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_enb
  )

# AMZN
remove_amzn <- c(
  "amazon", "amzn", "apple", "article", "analysis", "story", "stock", "stocks", "share", "price", "current", "investors", "company", "companies", "market", "narrative", "narratives", "based", "time", "analysis", "article", "story", "wall", "simply", "free", "top", "community", "globalstar", "term", "business", "analyst", "alternatively", "billions", "millions", "fair", "services", "buy", "potential", "valuation", "capital", "stake", "shares", "return", "retail", "margins", "financial", "deal", "data", "control", "support", "report", "past", "news", "nasdaq", "million", "dcf", "existing", "iphone", "satellite", "aws", "cloud", "worth", "ratio", "position", "future", "industry", "emergency", "recent", "quarter", "kuper", "portfolio", "forward", "head", "compared", "advertising", "moves", "billionaire", "businesses", "commentary", "directly", "estimate", "driven", "note", "infrastructure", "kuiper", "sits", "margins", "focused", "key", "discover", "platform", "reported", "features", "include", "sell", "investment", "tech", "power", "partnerships"
  )

remove_bigram_amzn <- c(
  "simply wall", "wall st", "data note", "cash flows", "current share", "share price", "nasdaq amzn", "million shares", "project kuiper", "historical data", "fundamental data", "financial situation", "financial advice", "team simplywallst", "research report", "qualitative material", "provide commentary", "email editorial", "editorial team", "stay updated", "sits alongside", "news stories", "unbiased methodology", "fair ratio", "multiline retail", "million individual", "past week", "nasdaqs aapl", "nasdaqs aapl", "nasdaqgs aapl", "iphone models", "assumes fair", "gains control", "low earth", "earth orbit"
  )

clean_amzn <- news_amzn %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::mutate(
    word = dplyr::case_when(
      word == "flows" ~ "flow",
      word == "risks" ~ "risk",
      word == "profit" ~ "earnings",
      TRUE ~ word
      )
    ) %>% 
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_amzn
  )

# LMT
remove_lmt <- c(
  "free", "community", "news", "said", "report", "reports", "according", "reuters", "article", "stock", "stocks", "buy", "hold", "investor", "investors", "canadian", "tfsa", "rrsp", "top", "best", "why", "keep", "eye", "built", "right", "now", "including", "based", "around", "well", "still", "also", "analysis", "story", "stock", "stocks", "share", "price", "current", "investors", "company", "companies", "market", "billions", "billion", "millions", "martin", "lockheed", "lmt", "trump", "nasa", "nyse", "artemis", "orion", "redwire", "production", "program", "programs", "system", "systems", "support", "term", "narrative", "increase", "space", "launch", "lunar", "zacks", "wall", "trillion", "simply", "moon", "president", "analysts", "crewed", "march", "center", "rtx", "worth", "past", "future", "operations", "data", "research", "key", "efforts", "expanding", "infrastructure", "financial", "human", "hardware", "spacecraft", "navigation", "discretionary", "critical", "exposure", "momentum", "airbus", "building", "businesses", "camera", "corporation", "capacity", "white", "portfolio", "noc", "iran", "include", "department", "directly", "house", "time", "successful", "world", "sun", "shares", "return", "rate", "yield", "sensors", "manufacturing", "integration", "imaging", "facility", "exploration", "friday", "government", "deep", "domestic"
  )

clean_lmt <- news_lmt %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::mutate(
    word = dplyr::case_when(
      word == "investments" ~ "investment",
      TRUE ~ word
      )
    ) %>%
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_lmt
  )

# BWXT
remove_bwxt <- c(
  "free", "community", "news", "said", "report", "reports", "according", "reuters", "article", "stock", "stocks", "buy", "hold", "investor", "investors", "canadian", "tfsa", "rrsp", "top", "best", "why", "keep", "eye", "built", "right", "now", "including", "based", "around", "well", "still", "also", "analysis", "story", "stock", "stocks", "share", "price", "current", "investors", "company", "companies", "market", "billions", "millions", "analysts", "bwx", "bwxt", "fair", "million", "zacks", "billion", "flr", "wall", "earth", "guidance", "commercial", "term", "simply", "clean", "days", "narrative", "projects", "business", "components", "services", "financial", "valuation", "estimate", "systems", "communication", "supply", "chain", "government", "time", "data", "contract", "publisher", "rank", "model", "access", "expected", "advanced", "results", "operations", "materials", "realoys", "oilprice", "technologies", "realaloys", "fluor", "contracts", "shares", "technology", "information", "project", "programs", "facility", "potential", "narratives", "global", "future", "analyst", "attention", "smr", "past", "program", "reactor", "propulsion", "visteon", "scale", "ceg", "ratio", "eps", "platforms", "facilities", "realaloys", "ceo", "capital", "constellation", "worth", "recent", "realloys", "consensus", "margins", "position", "markets"
  )

clean_bwxt <- news_bwxt %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>%
  tidytext::unnest_tokens(output = word, input = text) %>% 
  dplyr::mutate(
    word = dplyr::case_when(
      word == "growing" ~ "growth",
      word == "contracts" ~ "contract",
      word == "narratives" ~ "narrative",
      word == "programs" ~ "program",
      TRUE ~ word
      )
    ) %>% 
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_bwxt
  )

# Market
remove_market <- c(
  "free", "community", "yahoo", "read", "click", "follow", "news", "stock", "stocks", "market", "markets", "price", "prices", "finance", "financial", "business", "company", "companies", "president", "age", "social", "security", "retirement", "whoop", "trump", "iran", "global", "million", "street", "month", "wrote", "world", "quarter", "musk", "billion", "ceo", "analyst", "spacex", "wall", "tesla", "start", "reporter", "middle", "hill", "wednesday", "america", "analysis", "americans", "said", "says", "told", "added", "noted", "according", "year", "week", "day", "time", "today", "last", "new", "data", "report", "note", "expected", "estimate", "forecast", "country", "countries", "global", "economy", "economic", "hamza", "ines", "ferre", "nike", "pvt", "strait", "hormuz", "east", "major", "recent", "public", "moving", "events", "biggest", "impact", "top", "cut", "hit", "rising", "ahmed", "depth", "benefit", "concerns", "gulf", "march", "tuesday", "delta", "airlines", "infrastructure", "product", "monthly", "months", "analysts", "aren", "senior", "expectations", "investors", "health", "weeks", "space", "investing", "round", "ipo"
  )

clean_market <- news_market %>%
  clean_text() %>% 
  dplyr::select(article_num, text) %>% 
  tidytext::unnest_tokens(output = word, input = text) %>%
  dplyr::mutate(
    word = dplyr::case_when(
      word == "growing" ~ "growth",
      word == "contracts" ~ "contract",
      word == "narratives" ~ "narrative",
      word == "programs" ~ "program",
      word == "money" ~ "capital",
      word == "tariff" ~ "tariffs",
      TRUE ~ word
      )
    ) %>%
  dplyr::anti_join(stop_words, by = c("word")) %>% 
  dplyr::filter(
    stringr::str_length(word) > 2,
    !word %in% remove_market
  )
```

```{r}
clean_market %>%
  dplyr::count(word, sort = TRUE) %>% 
  dplyr::top_n(20) %>% 
  ggplot2::ggplot(ggplot2::aes(x = stats::reorder(word, n), y = n, fill = word)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::coord_flip() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::ggtitle("Most Used Words")
```

```{r, warning=FALSE}
sentiment_analysis <- function(data){
  # The Loughran Sentiment
  loughran <- data %>% 
    dplyr::inner_join(loughran, by = "word") %>% 
    dplyr::rename(loughran = sentiment) %>% 
    dplyr::count(article_num, loughran) %>% 
    tidyr::pivot_wider(
      names_from = loughran,
      values_from = n,
      values_fill = 0
    ) %>% 
    dplyr::mutate(net_loughran = positive - negative)
  
  loughran_score <- sum(loughran$net_loughran) / nrow(data)
  
  # The AFINN Sentiment
  afinn <- data %>% 
    dplyr::inner_join(afinn, by = "word") %>%
    dplyr::rename(net_afinn = value)
  
  afinn_score <- sum(afinn$net_afinn) / nrow(data)
  
  # The NRC Sentiment
  nrc <- data %>% 
    dplyr::inner_join(nrc, by = "word") %>% 
    dplyr::rename(nrc = sentiment) %>% 
    dplyr::count(article_num, nrc) %>% 
    tidyr::pivot_wider(
      names_from = nrc,
      values_from = n,
      values_fill = 0
    ) %>% 
    dplyr::mutate(total = positive + joy + trust + anticipation - fear - sadness - anger - disgust - negative)
  
  nrc_score <- sum(nrc$total) / nrow(data)
  
  # Calculation
  score <- 0.85 * loughran_score + 0.1 * afinn_score + 0.05 * nrc_score

  interpretation <- dplyr::case_when(
    score >= 0.025981 ~ "Strong Positive", # Above the 3rd quartile
    score >= 0.012856 ~ "Positive", # Between the Median and 3rd qurtile
    score >= 0.004982 ~ "Mild Positive", # Between the 1st quartile and the Median
    TRUE ~ "Weak Positive"
  )
  
  return(list(score = score, inter = interpretation))
}
```

```{r, warning=FALSE}
sentiment_analysis(clean_enb)


scores <- c(
  sentiment_analysis(clean_tsla)$score,
  sentiment_analysis(clean_enb)$score,
  sentiment_analysis(clean_amzn)$score,
  sentiment_analysis(clean_lmt)$score,
  sentiment_analysis(clean_bwxt)$score
)

summary(scores)
```

```{r}
news_amzn %>% 
  clean_text() %>% 
  tidytext::unnest_tokens(
    output = word,
    input = text,
    token = "ngrams",
    n = 2
  ) %>%
  tidyr::separate(word, c("w1", "w2"), sep = " ") %>% 
  dplyr::filter(
    !w1 %in% stop_words$word,
    !w2 %in% stop_words$word,
    !w1 %in% remove_amzn,
    !w2 %in% remove_amzn,
    nchar(w1) > 2,
    nchar(w2) > 2
    ) %>% 
  stats::na.omit() %>% 
  tidyr::unite(word, w1, w2, sep = " ") %>% 
  dplyr::filter(!word %in% remove_bigram_amzn) %>% 
  dplyr::count(word, sort = TRUE) %>% 
  dplyr::top_n(5) %>% 
  ggplot2::ggplot(ggplot2::aes(x = stats::reorder(word, n), y = n, fill = word)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::coord_flip() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::ggtitle("Most Used Words")
```

```{r}
assessment_stocks <- c("TSLA", "ENB", "AMZN", "LMT", "BWXT")
```

```{r, warning=FALSE}
# Grabing the assessment equity data
assessment <- tidyquant::tq_get(
  assessment_stocks,
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na()

assessment_wide <- assessment %>% 
  dplyr::select(date, symbol, return_log) %>% 
  tidyr::pivot_wider(names_from = symbol, values_from = return_log)


data_vix <- tidyquant::tq_get(
  "^VIX",
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(vix_return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na() %>% 
  dplyr::rename(vix_adjusted = adjusted) %>%
  dplyr::ungroup() %>% 
  dplyr::select(date, vix_adjusted, vix_return_log)

data_gspc <- tidyquant::tq_get(
  "^GSPC",
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(gspc_return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na() %>% 
  dplyr::rename(gspc_adjusted = adjusted) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(date, gspc_adjusted, gspc_return_log)
```

```{r}
technicals <- function(ticker, original = assessment, original_wide = assessment_wide, market_gspc = data_gspc, market_vix = data_vix){
  
  data_stock <- original %>% 
    dplyr::filter(symbol == ticker) %>% 
    dplyr::ungroup() %>% 
    dplyr::arrange(date)
  
  # MACD
  table <- TTR::MACD(
    original_wide[[ticker]],
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    maType = "EMA",
    percent = FALSE
    )
  
  table <- data_stock %>% 
    dplyr::select(date, open, close, high, low, volume, return_log) %>% 
    dplyr::left_join(
      as.data.frame(table) %>%  # Converting back to a data frame table
        dplyr::transmute(date = original_wide$date, macd, signal),
      by = "date"
      ) %>% 
    dplyr::mutate(macd_hist = macd - signal)
  
  
  # MFI
  xts_tbl <- xts::xts(
    x = data_stock %>% dplyr::select(open, high, low, close, volume),
    order.by = data_stock$date
    )

  tech_mfi <- TTR::MFI(
    HLC = xts_tbl[, c("high", "low", "close")],
    volume = xts_tbl[, "volume"],
    n = 14 # Using a 2 week period
    )

  tbl_mfi <- data.frame( 
    date = zoo::index(tech_mfi),
    zoo::coredata(tech_mfi)
    )

  table <- table %>% dplyr::left_join(tbl_mfi, by = "date")
  
  
  # Rolling SD
  table <- table %>% 
    dplyr::mutate(
      sd12 = slider::slide_dbl(
        .x = return_log,
        .f = sd,
        .before = 12, # Looking 3 months back
        .after = 0,
        .complete = FALSE 
        ) * sqrt(52)
      ) # I didn't remove NAs b/c other technicals like MACD need long beg. periods, so it will remove a lot of data
  
  
  # ^VIX and ^GSPC
  table <- table %>% 
    dplyr::left_join(market_vix, by = "date") %>% 
    dplyr::left_join(market_gspc, by = "date")
  
  
  
  # Plotting the data
  plot_return <- table %>% 
    plotly::plot_ly(
      x = ~date,
      y = ~return_log,
      type = "scatter",
      mode = "lines"
    ) %>% 
    plotly::style(name = "Log Return") # For the trace legend when using subplot

  plot_macd <- table %>%
    plotly::plot_ly(x = ~date) %>% 
    plotly::add_lines(x = ~date, y = ~macd, name = "MACD") %>% 
    plotly::add_lines(x = ~date, y = ~signal, name = "Signal")
  
  plot_sd12 <- table %>%
    plotly::plot_ly(
      x = ~date,
      y = ~sd12,
      type = "scatter",
      mode = "lines"
      ) %>%  
    plotly::style(name = "Rolling SD (12)")
  
  plot_mfi <- table %>% 
    plotly::plot_ly(
      x = ~date,
      y = ~mfi,
      type = "scatter",
      mode = "lines"
    ) %>%  
    plotly::style(name = "MFI")
  
  plot_mkt <- table %>% 
    plotly::plot_ly(x = ~date) %>%
    plotly::add_lines(x = ~date, y = ~return_log, name = "Log Returns") %>% 
    plotly::add_lines(x = ~date, y = ~gspc_return_log, name = "S&P 500")
  
  plot_vix <- table %>% 
    plotly::plot_ly(
      x = ~date,
      y = ~vix_adjusted,
      type = "scatter",
      mode = "lines"
    ) %>%  
    plotly::style(name = "VIX")
  
  
  plot_all_tech <- plotly::subplot(
    plot_return,
    plot_macd,
    plot_sd12,
    plot_mfi,
    nrows = 4,
    heights = c(0.30, 0.30, 0.20, 0.20), # Top plot get 25% of height, 2nd 25%, 3rd 25%, and bottom plot gets 25%
    shareX = TRUE,
    titleX = TRUE
    ) %>% 
    plotly::layout(
      title = list(text = paste(ticker, "Time Series by Week")),
      xaxis = list(title = "Date"),
      yaxis = list(title = "Log Return", titlefont = list(size = 12)),
      yaxis2 = list(title = "EMA", titlefont = list(size = 10)),
      yaxis3 = list(title = "Standard Deviation", titlefont = list(size = 10)),
      yaxis4 = list(title = "Momentum", titlefont = list(size = 10))
      )
  
  
  plot_all_mkt <- plotly::subplot(
    plot_mkt,
    plot_vix,
    nrows = 2,
    heights = c(0.55, 0.45),
    shareX = TRUE,
    titleX = TRUE
    ) %>% 
    plotly::layout(
      title = list(text = paste(ticker, "Time Series by Week")),
      xaxis = list(title = "Date"),
      yaxis = list(title = "Log Return", titlefont = list(size = 12)),
      yaxis2 = list(title = "Value", titlefont = list(size = 10))
      )
  
  return(list(table = table, plot_stock = plot_all_tech, plot_market = plot_all_mkt))
}
```

```{r}
x <- technicals("LMT")

x$plot_stock
```

```{r}
x$plot_stock
```

```{r}
x$plot_market
```

# The Analysis

## Tesla

## Enbridge

## Amazon

## Lockheed Martin

## BWX Technologies

# Conclusion

1.  Predictive modeling
    -   Future returns

    -   Probability of outperforming
2.  Financial news sentiment
    -   Does news impact returns?
3.  Risk / diversification
    -   Correlation with portfolio

    -   Volatility impact
4.  Feature relationships
    -   Non-linear effects (spilnes)

    -   Seasonality?

    -   PCA for dimentionality reduction

When adding the new company, I want to increase the expected return of the portfolio and decrease the risk (SD) as much as possible, but it's okay if it increases a little bit if the return increases a lot more.

Check for correlation of the company's return with the portfolio's return.

-   The smaller the greater risk reduction

We should choose the portfolio by seeing if its CAL is the steepest, so using the sharpe ratio which is the slope equation for that line.

To incorporate the financial news sentiment we can convert text news to numbers and then define what's good news what's bad news take the proportion of each one and use them as predictors for the model
