---
title: "Equity Portfolio Decision Making"
author: Daniel Rabi
date: April 9, 2026
backgroundcolor: lightblue
format: html
editor: visual
embed-resources: true
bibliography: references/references.bib
csl: references/apa.csl
output-file: index
nocite: "@YahooFinance, @Gratton_2025, @Dolan_2025, @Kuepper_2025"

echo: FALSE
message: FALSE
warning: FALSE
---

![[@Zerdzicki_2026]](images/into_picture.jpg){fig-align="center" width="600"}

Have you ever wondered how to decide whether you should add equity to your existing portfolio?

In this project, I will walk you through the decision-making process for assessing equities before adding them to an existing portfolio.

# The Portfolio

For this project's portfolio, I used a fixed, fairly diversified existing portfolio that contains the following companies:

```{r, message=FALSE, warning=FALSE}
library(gt)

dplyr::tibble(
  Equity = c(
    "NVIDIA",
    "Boeing",
    "Chevron",
    "Coca-Cola",
    "JPMorgan Chase & Co"
  ),
  Ticker = c(
    "NVDA",
    "BA",
    "CVX",
    "KO",
    "JPM"
  ),
  Description = c(
    "NVIDIA sells products in the Compute & Networking and Graphics segments with a high focus on automation and AI integration",
    "Boeing manufactures and sells products and services in segments like Commercial Airplanes, Defence, and Space & Security",
    "Chevron operates in all operational segments that provide energy resources across the world",
    "Coca-Cola manufactures and sells non-alcoholic beverages worldwide",
    "JPMorgan is a Commercial, Investment, and Asset & Wealth Management bank with operations around the world"
  ),
  Industry = c(
    "Technology industry",
    "Industrial industry",
    "Energy industry",
    "Beverages industry",
    "Finance services industry"
  ),
  Shares = c(
    "200",
    "200",
    "200",
    "200",
    "200"
  ),
  Weigth = c(
    "20%",
    "20%",
    "20%",
    "20%",
    "20%"
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Portfolio Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  ) %>% 
  gt::tab_source_note(source_note = "(Yahoo Finance, n.d.)")
```

## The Portfolio's Efficient Frontier

The efficient frontier is the set of portfolios plotted against their levels of risk and expected return [@Ganti_2025]. Therefore, @Ganti_2025 suggests that a portfolio is efficient if it provides the highest return for a given level of risk. However, to find the optimal portfolio, we need to identify the line with the highest Sharpe Ratio (slope) on the efficient frontier [@Vipond_2020].

The Sharpe Ratio equation:

$$
\text{Sharpe Ratio} = \frac{E(R_P) - r_f}{\sigma(R_P)}
$$

Where:

$r_f =$ The 10-year U.S. Treasury Bond [@FRED_2026]

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyquant)
library(lubridate)
library(rvest)
library(ggrepel)
library(TTR)
library(plotly)
library(xts)
library(zoo)
library(slider)
library(jsonlite)
```

```{r}
portfolio_stocks <- c("NVDA", "BA", "CVX", "KO", "JPM")


portfolio <- tidyquant::tq_get(
  portfolio_stocks,
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly # Converting to weekly data because I don't want the data to be too noisy but I need large enough sample size for ML
  ) %>% 
  dplyr::mutate(return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na()


portfolio_wide <- portfolio %>% 
  dplyr::select(date, symbol, return_log) %>% 
  tidyr::pivot_wider(names_from = symbol, values_from = return_log) 

portfolio_wide <- portfolio_wide %>% 
  dplyr::mutate(Portfolio_Return = 0.2 * portfolio_wide$NVDA + 0.2 * portfolio_wide$BA + 0.2 * portfolio_wide$CVX + 0.2 * portfolio_wide$KO + 0.2 * portfolio_wide$JPM)


returns <- portfolio_wide %>% 
  dplyr::select(-date, - Portfolio_Return)

ann_returns <- colMeans(returns) * 52
ann_cov <- stats::cov(returns, use = "pairwise.complete.obs") * 52


# Static webscraping for the Risk Free Rate

url <- "https://fred.stlouisfed.org/series/DGS10"

rf_rate <- rvest::read_html(url) %>% 
  rvest::html_element(".series-meta-observation-value") %>% # I'm using the . at the beginning b/c it's a class
  rvest::html_text(trim = TRUE) %>% 
  as.numeric()

rf_rate <- rf_rate / 100

# Grabbing the Risk Free Rate Using FRED API

r_f <- tidyquant::tq_get(
  "DGS10",
  get = "economic.data",
  from = Sys.Date() - 7,
  to = Sys.Date()
) %>% 
  dplyr::arrange(date)

r_f <- r_f[1, 3] %>% dplyr::pull(price)
r_f <- r_f / 100 # I want the Risk Free to match with the rest of the return values
```

```{r}
# Simulating different portfolio possibilities for the Efficient Frontier

set.seed(123) # For replicability

simulation <- replicate(5000, {
  rnd_vals <- runif(5) # Randomizing 5 numbers
  weights <- rnd_vals / sum(rnd_vals) # Creating weights that sum up to 1
  
  port_returns <- sum(weights * ann_returns)
  port_sd <- sqrt(sum(weights * (ann_cov %*% weights))) # The %*% is for multiplying matrecies together
  
  c(
    NVDA = weights[1],
    BA = weights[2],
    CVX = weights[3],
    KO = weights[4],
    JPM = weights[5],
    return = port_returns,
    sd = as.numeric(port_sd) # as.numeric is needed b/c we used a matrix multiplication
  )
})

simulation <- as.data.frame(t(simulation)) # The t transposes, or flips, the rows and columns


simulation <- simulation %>% 
  dplyr::mutate(sharpe_ratio = (return - r_f) / sd) # Calculating the sharpe ratio to every portfolio simulation

best_sharpe_ratio <- simulation %>% 
  dplyr::slice_max(order_by = sharpe_ratio, n = 1)

mvp <- simulation %>% 
  dplyr::slice_min(order_by = sd, n = 1) # Keeping the highest values of SD in every equity

portfolio_exp_return <- mean(portfolio_wide$Portfolio_Return) * 52
portfolio_exp_sd <- sd(portfolio_wide$Portfolio_Return) * sqrt(52)
sharpe_ratio <- (portfolio_exp_return - r_f) / portfolio_exp_sd


optimal_table <- dplyr::tibble(
  sd = seq(from = 0, to = max(simulation$sd), length.out = 100),
  return = best_sharpe_ratio$sharpe_ratio * sd + r_f
)

portfolio_table <- dplyr::tibble(
  sd = portfolio_exp_sd,
  return = portfolio_exp_return
)


simulation %>% ggplot2::ggplot(ggplot2::aes(x = sd, y = return, color = return)) +
  ggplot2::geom_point(alpha = 0.3) + 
  ggplot2::scale_color_viridis_c() +
  ggplot2::geom_point(
    data = best_sharpe_ratio,
    ggplot2::aes(x = sd, y = return),
    color = "orange",
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = best_sharpe_ratio,
    ggplot2::aes(x = sd, y = return, label = "Optimal Portfolio"),
    color = "black"
  ) +
  ggplot2::geom_line(
    data = optimal_table,
    ggplot2::aes(x = sd, y = return),
    color = "orange"
  ) +
  ggplot2::geom_point(
    data = portfolio_table,
    ggplot2::aes(x = portfolio_exp_sd, y = portfolio_exp_return),
    color = "palevioletred1",
    size = 2
  ) + 
  ggrepel::geom_text_repel(
    data = portfolio_table,
    ggplot2::aes(x = portfolio_exp_sd, y = portfolio_exp_return, label = "Current Portfolio"),
    color = "black"
  ) +
  ggplot2::labs(
    title = "The Efficient Frontier",
    x = "Annualized Standard Deviation (Risk)",
    y = "Annualized Expected Log Return",
    color = "Log Return"
  ) +
  ggplot2::theme(plot.title = element_text(hjust = 0.5))
```

# The List of Potential Equities

The five equities that I have decided to assess are in the following table:

```{r}
dplyr::tibble(
  Equity = c(
    "Tesla",
    "Enbridge",
    "Amazon",
    "Lockheed Martin Corp",
    "BWX Technologies Inc."
  ),
  Ticker = c(
    "TSLA",
    "ENB",
    "AMZN",
    "LMT",
    "BWXT"
  ),
  Description = c(
    "Tesla manufactures and sells electric vehicles, along with energy generation and storage systems, worldwide",
    "Enbridge is a Canadian company that provides energy resources to both the US and Canada to power their operations",
    "Amazon is an online retail store with some physical stores worldwide, and an expansion into database services",
    "Lockheed Martin is an aerospace and defence company that manufactures and sells weapons and vehicles worldwide",
    "BWX manufactures and sells nuclear components worldwide, both to governments and commercial operations"
  ),
  Industry = c(
    "Technology & Automotive industry",
    "Energy industry",
    "Technology & Consumer industry",
    "Defense industry",
    "Nuclear & Defence industry"
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Potential Equities Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  ) %>% 
  gt::tab_source_note(source_note = "(Yahoo Finance, n.d.)")
```

# The Predictors

To fairly assess each equity, I aim to estimate where each equity is heading one week ahead. Therefore, I use the following predictors to estimate next week's log returns for each equity.

It is important to note that, for sentiment analysis, I selected specific lexicons for each equity, since they are scattered across different industries and sectors, which means that the same news could be long for one and short for the other.

```{r}
dplyr::tibble(
  Predictor = c(
    "Historical Log Returns",
    "MACD",
    "MFI",
    "Rolling SD (12)",
    "S&P 500",
    "VIX",
    "Positive News"
  ),
  Classification = c(
    "Technicals",
    "Technicals",
    "Technicals",
    "Technicals",
    "Market",
    "Market",
    "Financial News Sentiment"
  ),
  Description = c(
    "Captures short-term momentum and trend of the time series",
    "The Moving Average Convergence Divergence helps identify and measure trends in price and momentum, and to identify entry and exit points (Dolan, 2025).",
    "Analyzes price movements and trading volumes to measure the flow of money into and out of an asset (Gratton, 2025).",
    "Using a moving standard deviation of the previous 12 weeks helps smooth out random spikes and capture short-term market transitions.",
    'Using "^GSPC" rather than "SPY" provides a direct comparison to the market, allowing capture of overall market movements and isolation of stock-specific returns.',
    "The Volatility Index, or the Fear Index, measures expected market volatility and is used to evaluate market sentiment and perceived risk (Kuepper, 2025).",
    "Measures the tone of financial news, where positive sentiment may signal investor optimism and future price increases by using specific lexicons for each equity."
  )
) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "The Predictors Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  )
```

```{r, eval=FALSE}
options(mc.cores = parallel::detectCores()) # Utilizing all the cores

# Dynamic Webscraping
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "C:/Program Files/nodejs", sep=";"))

# TSLA News
out_tsla <- system("node ./scripts/TSLA/news_tsla.js", intern = TRUE)
txt_tsla <- paste(readLines("./scripts/TSLA/articles_tsla.json", warn = FALSE), collapse = "\n")

# ENB News
out_enb <- system("node ./scripts/ENB/news_enb.js", intern = TRUE)
txt_enb <- paste(readLines("./scripts/ENB/articles_enb.json", warn = FALSE), collapse = "\n")

# AMZN News
out_amzn <- system("node ./scripts/AMZN/news_amzn.js", intern = TRUE)
txt_amzn <- paste(readLines("./scripts/AMZN/articles_amzn.json", warn = FALSE), collapse = "\n")

# LMT News
out_lmt <- system("node ./scripts/LMT/news_lmt.js", intern = TRUE)
txt_lmt <- paste(readLines("./scripts/LMT/articles_lmt.json", warn = FALSE), collapse = "\n")

# BWXT News
out_bwxt <- system("node ./scripts/BWXT/news_bwxt.js", intern = TRUE)
txt_bwxt <- paste(readLines("./scripts/BWXT/articles_bwxt.json", warn = FALSE), collapse = "\n")

# Market News
out_market <- system("node ./scripts/Market/news_market.js", intern = TRUE)
txt_market <- paste(readLines("./scripts/Market/articles_market.json", warn = FALSE), collapse = "\n")
```

```{r}
# Assigning scraped text from the saved JSON file

# TSLA News
news_tsla <- jsonlite::fromJSON(txt_tsla)

# ENB News
news_enb <- jsonlite::fromJSON(txt_enb)

# AMZN News
news_amzn <- jsonlite::fromJSON(txt_amzn)

# LMT News
news_lmt <- jsonlite::fromJSON(txt_lmt)

# BWXT News
news_bwxt <- jsonlite::fromJSON(txt_bwxt)

# Market News
news_market <- jsonlite::fromJSON(txt_market)
```

```{r}
assessment_stocks <- c("TSLA", "ENB", "AMZN", "LMT", "BWXT")
```

```{r, warning=FALSE}
# Grabing the assessment equity data
assessment <- tidyquant::tq_get(
  assessment_stocks,
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na()

assessment_wide <- assessment %>% 
  dplyr::select(date, symbol, return_log) %>% 
  tidyr::pivot_wider(names_from = symbol, values_from = return_log)


data_vix <- tidyquant::tq_get(
  "^VIX",
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(vix_return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na() %>% 
  dplyr::rename(vix_adjusted = adjusted) %>%
  dplyr::ungroup() %>% 
  dplyr::select(date, vix_adjusted, vix_return_log)

data_gspc <- tidyquant::tq_get(
  "^GSPC",
  from = Sys.Date() - lubridate::years(5),
  to = Sys.Date(),
  get = "stock.prices"
) %>% 
  dplyr::group_by(symbol) %>% 
  tidyquant::tq_transmute(
    mutate_fun = to.weekly
  ) %>% 
  dplyr::mutate(gspc_return_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  tidyr::drop_na() %>% 
  dplyr::rename(gspc_adjusted = adjusted) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(date, gspc_adjusted, gspc_return_log)
```

```{r}
technicals <- function(ticker, original = assessment, original_wide = assessment_wide, market_gspc = data_gspc, market_vix = data_vix){
  
  data_stock <- original %>% 
    dplyr::filter(symbol == ticker) %>% 
    dplyr::ungroup() %>% 
    dplyr::arrange(date)
  
  # MACD
  table <- TTR::MACD(
    original_wide[[ticker]],
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    maType = "EMA",
    percent = FALSE
    )
  
  table <- data_stock %>% 
    dplyr::select(date, open, close, high, low, volume, return_log) %>% 
    dplyr::left_join(
      as.data.frame(table) %>%  # Converting back to a data frame table
        dplyr::transmute(date = original_wide$date, macd, signal),
      by = "date"
      ) %>% 
    dplyr::mutate(macd_hist = macd - signal)
  
  
  # MFI
  xts_tbl <- xts::xts(
    x = data_stock %>% dplyr::select(open, high, low, close, volume),
    order.by = data_stock$date
    )

  tech_mfi <- TTR::MFI(
    HLC = xts_tbl[, c("high", "low", "close")],
    volume = xts_tbl[, "volume"],
    n = 14 # Using a 2 week period
    )

  tbl_mfi <- data.frame( 
    date = zoo::index(tech_mfi),
    zoo::coredata(tech_mfi)
    )

  table <- table %>% dplyr::left_join(tbl_mfi, by = "date")
  
  
  # Rolling SD
  table <- table %>% 
    dplyr::mutate(
      sd12 = slider::slide_dbl(
        .x = return_log,
        .f = sd,
        .before = 12, # Looking 3 months back
        .after = 0,
        .complete = FALSE 
        ) * sqrt(52)
      ) # I didn't remove NAs b/c other technicals like MACD need long beg. periods, so it will remove a lot of data
  
  
  # ^VIX and ^GSPC
  table <- table %>% 
    dplyr::left_join(market_vix, by = "date") %>% 
    dplyr::left_join(market_gspc, by = "date")
  
  
  
  # Plotting the data
  plot_return <- table %>% 
    plotly::plot_ly(
      x = ~date,
      y = ~return_log,
      type = "scatter",
      mode = "lines"
    ) %>% 
    plotly::style(name = "Log Return") # For the trace legend when using subplot

  plot_macd <- table %>%
    plotly::plot_ly(x = ~date) %>% 
    plotly::add_lines(x = ~date, y = ~macd, name = "MACD") %>% 
    plotly::add_lines(x = ~date, y = ~signal, name = "Signal")
  
  plot_sd12 <- table %>%
    plotly::plot_ly(
      x = ~date,
      y = ~sd12,
      type = "scatter",
      mode = "lines"
      ) %>%  
    plotly::style(name = "Rolling SD (12)")
  
  plot_mfi <- table %>% 
    plotly::plot_ly(
      x = ~date,
      y = ~mfi,
      type = "scatter",
      mode = "lines"
    ) %>%  
    plotly::style(name = "MFI")
  
  plot_mkt <- table %>% 
    plotly::plot_ly(x = ~date) %>%
    plotly::add_lines(x = ~date, y = ~return_log, name = "Log Returns") %>% 
    plotly::add_lines(x = ~date, y = ~gspc_return_log, name = "S&P 500")
  
  plot_vix <- table %>% 
    plotly::plot_ly(
      x = ~date,
      y = ~vix_adjusted,
      type = "scatter",
      mode = "lines"
    ) %>%  
    plotly::style(name = "VIX")
  
  
  plot_all_tech <- plotly::subplot(
    plot_return,
    plot_macd,
    plot_sd12,
    plot_mfi,
    nrows = 4,
    heights = c(0.30, 0.30, 0.20, 0.20), # Top plot get 25% of height, 2nd 25%, 3rd 25%, and bottom plot gets 25%
    shareX = TRUE,
    titleX = TRUE
    ) %>% 
    plotly::layout(
      title = list(text = paste(ticker, "Time Series by Week")),
      xaxis = list(title = "Date"),
      yaxis = list(title = "Log Return", titlefont = list(size = 12)),
      yaxis2 = list(title = "EMA", titlefont = list(size = 10)),
      yaxis3 = list(title = "Standard Deviation", titlefont = list(size = 10)),
      yaxis4 = list(title = "Momentum", titlefont = list(size = 10))
      )
  
  
  plot_all_mkt <- plotly::subplot(
    plot_mkt,
    plot_vix,
    nrows = 2,
    heights = c(0.55, 0.45),
    shareX = TRUE,
    titleX = TRUE
    ) %>% 
    plotly::layout(
      title = list(text = paste(ticker, "Time Series by Week")),
      xaxis = list(title = "Date"),
      yaxis = list(title = "Log Return", titlefont = list(size = 12)),
      yaxis2 = list(title = "Value", titlefont = list(size = 10))
      )
  
  return(list(table = table, plot_stock = plot_all_tech, plot_market = plot_all_mkt))
}
```

```{r}
x <- technicals("LMT")

x$plot_stock
```

```{r}
x$plot_stock
```

```{r}
x$plot_market
```

1.  Predictive modeling
    -   Future returns

    -   Probability of outperforming
2.  Financial news sentiment
    -   Does news impact returns?
3.  Risk / diversification
    -   Correlation with portfolio

    -   Volatility impact
4.  Feature relationships
    -   Non-linear effects (spilnes)

    -   Seasonality?

    -   PCA for dimentionality reduction

When adding the new company, I want to increase the expected return of the portfolio and decrease the risk (SD) as much as possible, but it's okay if it increases a little bit if the return increases a lot more.

Check for correlation of the company's return with the portfolio's return.

-   The smaller the greater risk reduction

We should choose the portfolio by seeing if its CAL is the steepest, so using the sharpe ratio which is the slope equation for that line.

To incorporate the financial news sentiment we can convert text news to numbers and then define what's good news what's bad news take the proportion of each one and use them as predictors for the model

# Predictors:

Technical:

-   Moving average (MACD)

-   MFI

-   Rolling volatility

-   Past returns 1 week

-   Past returns 4 week

Market:

-   Market return

-   VIX

Financial News:

-   Avg. news 7 days

-   Avg. news 30 days

-   Headline count 7 days

-   Negative ratio 7 days

I need to scrape the risk free rate for the Capital Market Line for the optimal portfolio
